---
title: "STA 9750 Mini-Project #01: Assessing the Impact of SFFA on Campus Diversity"
author: "Adnan Sarker"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    code-summary: "Show Code"
    theme: cosmo
    embed-resources: true
execute:
  warning: false
  message: false
  cache: true
---

## Introduction

On June 29, 2023, the U.S. Supreme Court issued its landmark ruling in *Students
for Fair Admissions v. Harvard* (SFFA), effectively ending race-conscious
admissions at American colleges and universities. For institutions that had long
relied on holistic, race-aware review to build diverse classes, the decision
raised an urgent empirical question: how much did these programs actually matter?

This report uses enrollment data from IPEDS (the Integrated Postsecondary
Education Data System) to measure changes in campus diversity before and after
SFFA. We focus on first-year undergraduate enrollment — the cohort most directly
affected by the ruling — and compare trends nationally, at CUNY institutions, and
against California's public universities, which have been barred from affirmative
action since 1996 and therefore serve as a useful baseline.

---

## Data Acquisition

```{r acquire-data}
#| label: acquire-data
#| cache: true

#' Acquire IPEDS Data for MP#01
#'
#' Downloads fall enrollment ("EF") and institution description ("HD") files
#' for each year in [start_year, end_year] from IPEDS 
acquire_ipeds_data <- function(start_year = 2010, end_year = 2024) {
  library(tidyverse)
  library(glue)

  data_dir <- file.path("data", "mp01")

  if (!dir.exists(data_dir)) {
    dir.create(data_dir, showWarnings = FALSE, recursive = TRUE)
  }

  YEARS <- seq(start_year, end_year)

  EFA_ALL <- map(YEARS, function(yy) {
    if (yy <= 2022) {
      ef_url <- glue("https://nces.ed.gov/ipeds/datacenter/data/EF{yy}A.zip")
    } else {
      ef_url <- glue(
        "https://nces.ed.gov/ipeds/data-generator?year={yy}",
        "&tableName=EF{yy}A&HasRV=0&type=csv"
      )
    }

    ef_file <- file.path(data_dir, glue("ef{yy}a.csv.zip"))

    if (!file.exists(ef_file)) {
      message(glue("Downloading Enrollment Data for {yy} from {ef_url}"))
      download.file(ef_url, destfile = ef_file, quiet = TRUE, mode = "wb")
    }

    read_csv(ef_file, show_col_types = FALSE) |>
      mutate(
        year = yy,
        enrollment_m_aian = EFAIANM,
        enrollment_f_aian = EFAIANW,
        enrollment_m_asia = EFASIAM,
        enrollment_f_asia = EFASIAW,
        enrollment_m_bkaa = EFBKAAM,
        enrollment_f_bkaa = EFBKAAW,
        enrollment_m_hisp = EFHISPM,
        enrollment_f_hisp = EFHISPW,
        enrollment_m_nhpi = EFNHPIM,
        enrollment_f_nhpi = EFNHPIW,
        enrollment_m_whit = EFWHITM,
        enrollment_f_whit = EFWHITW,
        enrollment_m_2mor = EF2MORM,
        enrollment_f_2mor = EF2MORW,
        enrollment_m_unkn = EFUNKNM,
        enrollment_f_unkn = EFUNKNW,
        enrollment_m_nral = EFNRALM,
        enrollment_f_nral = EFNRALW
      ) |>
      filter(
        (EFALEVEL %in% c(2, 12)) | (LINE %in% c(1, 15))
      ) |>
      mutate(level = case_when(
        EFALEVEL == 2          ~ "all undergrad",
        EFALEVEL == 12         ~ "all graduate",
        LINE %in% c(1, 15)     ~ "first year undergrad"
      )) |>
      select(
        institution_id = UNITID,
        year,
        level,
        starts_with("enrollment_")
      ) |>
      group_by(institution_id, year, level) |>
      summarize(across(starts_with("enrollment_"), sum), .groups = "drop")
  }) |> bind_rows()

  DESC_ALL <- map(YEARS, function(yy) {
    if (yy <= 2022) {
      hd_url <- glue("https://nces.ed.gov/ipeds/datacenter/data/HD{yy}.zip")
    } else {
      hd_url <- glue(
        "https://nces.ed.gov/ipeds/data-generator?year={yy}",
        "&tableName=HD{yy}&HasRV=0&type=csv"
      )
    }

    hd_file <- file.path(data_dir, glue("hd{yy}.csv.zip"))

    if (!file.exists(hd_file)) {
      message(glue("Downloading Institutional Descriptions for {yy} from {hd_url}"))
      download.file(hd_url, destfile = hd_file, quiet = TRUE, mode = "wb")
    }

    suppressWarnings(
      read_csv(
        hd_file,
        show_col_types = FALSE,
        locale = locale(encoding = if_else(yy == 2024, "utf-8", "windows-1252"))
      ) |>
        mutate(year = yy) |>
        select(
          institution_id = UNITID,
          institution_name = INSTNM,
          state = STABBR,
          year
        )
    )
  }) |> bind_rows()

  inner_join(EFA_ALL, DESC_ALL,
    join_by(institution_id == institution_id, year == year)
  )
}

IPEDS <- acquire_ipeds_data()
```

The call above ^ downloads 15 years (2010–2024) of IPEDS fall-enrollment and
institutional-description data, caching under `data/mp01/`. The resulting data frame contains
`r nrow(IPEDS) |> scales::comma()` rows and `r ncol(IPEDS)` columns.

---

## Data Cleaning and Preparation

### Task 2: Identify CUNY Schools

We add a Boolean column `is_cuny` that is `TRUE` for any institution whose name
contains the substring `"CUNY"`.

```{r cuny-flag}
library(tidyverse)
library(scales)

IPEDS <- IPEDS |>
  mutate(is_cuny = str_detect(institution_name, "CUNY"))

# Quick check — how many distinct CUNY institutions?
n_cuny <- IPEDS |> filter(is_cuny) |> distinct(institution_id) |> nrow()
```

`str_detect` scans each institution name for the substring `"CUNY"` and returns
`TRUE` when found. The mutate assigns this vector as a new column. There are
`r n_cuny` distinct CUNY institutions in the data set.

### Task 3: Identify UC and Cal State Schools

California banned affirmative action in public higher education via Proposition
209 (1996), long before SFFA. Its public universities therefore serve as a
natural control group: any demographic shift we observe there after 2023 cannot
be attributed to the ruling.

We identify the two selective systems — the University of California and
California State — using two separate `str_detect` calls combined with the `|`
operator. Community colleges are excluded because they are open-enrollment.

```{r calpublic-flag}
IPEDS <- IPEDS |>
  mutate(
    is_calpublic =
      str_detect(institution_name, "University of California") |
      str_detect(institution_name, "California State University")
  )

# Verify: list all identified CA public institutions
ca_schools <- IPEDS |>
  filter(is_calpublic) |>
  distinct(institution_name) |>
  arrange(institution_name)
```

The logic flags `TRUE` when an institution name contains either
`"University of California"` *or* `"California State University"`. Reviewing the
`r nrow(ca_schools)` distinct campuses identified confirms the list matches
the UC and CSU systems.

---

## Initial Data Exploration

### Task 4: Initial EDA Pass

Before any substantive analysis, we check data types and basic summary
statistics — two habits that catch silent errors early.

**Type check with `glimpse()`:**

```{r glimpse-data}
glimpse(IPEDS)
```

All 18 enrollment columns are numeric (`<dbl>`), as expected. `institution_name`
and `state` are character strings; `year` and `level` carry the year integer and
level label respectively; and our two new Boolean columns (`is_cuny`,
`is_calpublic`) are logical. No type surprises.

**Summary statistics with `skimr::skim()`:**

```{r skim-data}
library(skimr)
skim(IPEDS)
```

A few things stand out from the skim output:

- Many enrollment cells are zero (or very small) —> this is expected, as not all
  institutions have graduate programs and small colleges will have zero students
  in rare demographic categories.
- Columns for `nhpi` (Native Hawaiian / Pacific Islander) show near-zero medians
  nationally, consistent with this group's small overall share of the
  undergraduate population.
- No column-level missingness is apparent in the numeric enrollment fields beyond
  what is expected from the data structure.

---

## Exploratory Analysis

Before computing anything, it is useful to add a few derived columns we will
need repeatedly: total enrollment, total female enrollment, and a
"non-white, non-Asian" diversity count.

```{r derived-columns}
IPEDS <- IPEDS |>
  mutate(
    # Total enrolled (sum of all 18 cells)
    total_enrolled = rowSums(across(starts_with("enrollment_"))),

    # Total female enrolled
    total_female = rowSums(across(starts_with("enrollment_f_"))),

    # Total male enrolled
    total_male = rowSums(across(starts_with("enrollment_m_"))),

    # White students (both sexes)
    total_white = enrollment_m_whit + enrollment_f_whit,

    # Asian students (both sexes)
    total_asian = enrollment_m_asia + enrollment_f_asia,

    # "URM" for this analysis: non-white AND non-Asian domestic students
    # (excludes Unknown and Non-Resident Alien as we can't classify them)
    total_urm = enrollment_m_aian + enrollment_f_aian +
                enrollment_m_bkaa + enrollment_f_bkaa +
                enrollment_m_hisp + enrollment_f_hisp +
                enrollment_m_nhpi + enrollment_f_nhpi +
                enrollment_m_2mor + enrollment_f_2mor,

    # Fraction female
    frac_female  = total_female  / total_enrolled,

    # Fraction white
    frac_white   = total_white   / total_enrolled,

    # Fraction URM
    frac_urm     = total_urm     / total_enrolled
  )
```

### Task 5: Inline Answers

**Q1 — How many distinct institutions appear in the data set?**

```{r q1}
n_institutions <- IPEDS |> distinct(institution_id) |> nrow()
```

There are **`r comma(n_institutions)`** distinct institutions in the IPEDS data.

---

**Q2 — Graduate students enrolled at Baruch in 2024?**

```{r q2}
baruch_grad_2024 <- IPEDS |>
  filter(
    str_detect(institution_name, "Baruch"),
    year == 2024,
    level == "all graduate"
  ) |>
  summarize(grad = sum(total_enrolled)) |>
  pull(grad)
```

In Fall 2024, **`r comma(baruch_grad_2024)`** graduate students were enrolled at
CUNY Bernard M. Baruch College.

---

**Q3 — Total students at Baruch in 2024 (no double-counting)?**

`"first year undergrad"` is a subset of `"all undergrad"`, so we sum only the
two non-overlapping levels: `"all undergrad"` and `"all graduate"`.

```{r q3}
baruch_total_2024 <- IPEDS |>
  filter(
    str_detect(institution_name, "Baruch"),
    year == 2024,
    level %in% c("all undergrad", "all graduate")
  ) |>
  summarize(total = sum(total_enrolled)) |>
  pull(total)
```

Total enrollment at Baruch in Fall 2024 was **`r comma(baruch_total_2024)`**
students.

---

**Q4 — Institution with the most enrolled female students in 2019?**

```{r q4}
most_female_2019 <- IPEDS |>
  filter(year == 2019, level == "all undergrad") |>
  slice_max(total_female, n = 1) |>
  select(institution_name, total_female)

q4_name  <- most_female_2019$institution_name
q4_count <- most_female_2019$total_female
```

In 2019, **`r q4_name`** had the highest number of enrolled female undergraduate
students, with **`r comma(q4_count)`** women.

---

**Q5 — Highest proportion of NHPI first-year undergrads in 2024 (>1,000 total)?**

```{r q5}
nhpi_top_2024 <- IPEDS |>
  filter(year == 2024, level == "first year undergrad") |>
  mutate(
    nhpi_count = enrollment_m_nhpi + enrollment_f_nhpi,
    nhpi_frac  = nhpi_count / total_enrolled
  ) |>
  filter(total_enrolled >= 1000) |>
  slice_max(nhpi_frac, n = 1) |>
  select(institution_name, nhpi_frac, nhpi_count, total_enrolled)

q5_name <- nhpi_top_2024$institution_name
q5_frac <- nhpi_top_2024$nhpi_frac
```

Among institutions enrolling at least 1,000 first-year undergraduates,
**`r q5_name`** had the highest fraction of Native Hawaiian / Pacific Islander
students in 2024, at **`r percent(q5_frac, accuracy = 0.1)`** of the entering
class.

---

### Task 6: Table Answers

```{r load-gt}
library(gt)
```

**T6-Q1 — Top 5 states by total graduate enrollment across all years**

```{r t6q1}
IPEDS |>
  filter(level == "all graduate") |>
  group_by(state) |>
  summarize(grad_total = sum(total_enrolled, na.rm = TRUE), .groups = "drop") |>
  slice_max(grad_total, n = 5) |>
  arrange(desc(grad_total)) |>
  gt() |>
  cols_label(state = "State", grad_total = "Total Graduate Enrollment") |>
  fmt_number(grad_total, decimals = 0) |>
  tab_header(
    title    = "Top 5 States by Graduate Student Enrollment",
    subtitle = "Aggregated across all institutions and years (2010–2024)"
  ) |>
  tab_source_note("Source: IPEDS via NCES")
```

---

**T6-Q2 — CUNY first-year undergraduate enrollment in 2024**

```{r t6q2}
cuny_total_fy <- IPEDS |>
  filter(is_cuny, year == 2024, level == "first year undergrad") |>
  summarize(system_total = sum(total_enrolled)) |>
  pull(system_total)

IPEDS |>
  filter(is_cuny, year == 2024, level == "first year undergrad") |>
  select(institution_name, total_enrolled) |>
  arrange(desc(total_enrolled)) |>
  mutate(pct_of_cuny = total_enrolled / cuny_total_fy) |>
  gt() |>
  cols_label(
    institution_name = "Institution",
    total_enrolled   = "First-Year Enrollment",
    pct_of_cuny      = "% of CUNY Total"
  ) |>
  fmt_number(total_enrolled, decimals = 0) |>
  fmt_percent(pct_of_cuny, decimals = 1) |>
  tab_header(
    title    = "CUNY First-Year Undergraduate Enrollment (Fall 2024)",
    subtitle = "By campus, with share of system-wide total"
  ) |>
  tab_source_note("Source: IPEDS via NCES") |>
  grand_summary_rows(
    columns  = total_enrolled,
    fns      = list(Total = ~ sum(.)),
    fmt      = ~ fmt_number(., decimals = 0)
  )
```

---

**T6-Q3 — Baruch total undergraduate enrollment over time**

```{r t6q3}
IPEDS |>
  filter(str_detect(institution_name, "Baruch"), level == "all undergrad") |>
  select(year, total_enrolled) |>
  arrange(year) |>
  mutate(yoy_pct = (total_enrolled - lag(total_enrolled)) / lag(total_enrolled)) |>
  gt() |>
  cols_label(
    year           = "Year",
    total_enrolled = "Undergraduate Enrollment",
    yoy_pct        = "YoY Change"
  ) |>
  fmt_number(total_enrolled, decimals = 0) |>
  fmt_percent(yoy_pct, decimals = 1) |>
  sub_missing(columns = yoy_pct, missing_text = "—") |>
  tab_header(
    title    = "Baruch College Undergraduate Enrollment (2010–2024)",
    subtitle = "Year-over-year percentage change"
  ) |>
  tab_source_note("Source: IPEDS via NCES")
```

---

**T6-Q4 — 5 institutions where the white student fraction fell most (2010–2020)**

We restrict to institutions with at least 1,000 undergraduates in both years to
avoid large swings driven by small-enrollment schools.

```{r t6q4}
white_wide <- IPEDS |>
  filter(level == "all undergrad", year %in% c(2010, 2020)) |>
  filter(total_enrolled >= 1000) |>
  select(institution_id, institution_name, year, frac_white) |>
  pivot_wider(
    names_from   = year,
    values_from  = frac_white,
    names_prefix = "frac_"
  ) |>
  drop_na() |>
  mutate(change = frac_2020 - frac_2010) |>
  slice_min(change, n = 5) |>
  select(institution_name, frac_2010, frac_2020, change) |>
  arrange(change)

white_wide |>
  gt() |>
  cols_label(
    institution_name = "Institution",
    frac_2010        = "White % (2010)",
    frac_2020        = "White % (2020)",
    change           = "Change (pp)"
  ) |>
  fmt_percent(c(frac_2010, frac_2020, change), decimals = 1) |>
  tab_header(
    title    = "Largest Declines in White Student Fraction (2010–2020)",
    subtitle = "Among institutions with ≥1,000 undergraduates in both years"
  ) |>
  tab_source_note("Source: IPEDS via NCES")
```

---

**T6-Q5 — 3 states with largest increase in female undergraduate fraction (2010–2024)**

```{r t6q5}
IPEDS |>
  filter(level == "all undergrad", year %in% c(2010, 2024)) |>
  group_by(state, year) |>
  summarize(
    female = sum(total_female,   na.rm = TRUE),
    total  = sum(total_enrolled, na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(frac_female = female / total) |>
  select(state, year, frac_female) |>
  pivot_wider(
    names_from   = year,
    values_from  = frac_female,
    names_prefix = "frac_"
  ) |>
  drop_na() |>
  mutate(change = frac_2024 - frac_2010) |>
  slice_max(change, n = 3) |>
  arrange(desc(change)) |>
  select(state, frac_2010, frac_2024, change) |>
  gt() |>
  cols_label(
    state     = "State",
    frac_2010 = "Female % (2010)",
    frac_2024 = "Female % (2024)",
    change    = "Change (pp)"
  ) |>
  fmt_percent(c(frac_2010, frac_2024, change), decimals = 1) |>
  tab_header(
    title    = "Largest Increases in Female Undergraduate Fraction by State",
    subtitle = "2010 vs. 2024"
  ) |>
  tab_source_note("Source: IPEDS via NCES")
```

---

## Op-Ed Analysis (Supporting Code)

The following code block computes all values referenced inline in the Op-Ed
below. It is hidden in the rendered document but fully reproducible.

```{r oped-calculations}
# ── Baruch identity ──────────────────────────────────────────────────────────
baruch_data <- IPEDS |>
  filter(str_detect(institution_name, "Baruch"))

# First-year class size: 2022, 2023, 2024
baruch_fy <- baruch_data |>
  filter(level == "first year undergrad", year %in% c(2022, 2023, 2024)) |>
  select(year, total_enrolled, total_urm, frac_urm, frac_female)

baruch_fy_2022 <- baruch_fy |> filter(year == 2022) |> pull(total_enrolled)
baruch_fy_2023 <- baruch_fy |> filter(year == 2023) |> pull(total_enrolled)
baruch_fy_2024 <- baruch_fy |> filter(year == 2024) |> pull(total_enrolled)

baruch_urm_frac_2023 <- baruch_fy |> filter(year == 2023) |> pull(frac_urm)
baruch_urm_frac_2024 <- baruch_fy |> filter(year == 2024) |> pull(frac_urm)
baruch_urm_change    <- baruch_urm_frac_2024 - baruch_urm_frac_2023

# All-undergrad URM fraction: 2023 vs 2024
baruch_ug <- baruch_data |>
  filter(level == "all undergrad", year %in% c(2023, 2024)) |>
  select(year, frac_urm)

baruch_ug_urm_2023 <- baruch_ug |> filter(year == 2023) |> pull(frac_urm)
baruch_ug_urm_2024 <- baruch_ug |> filter(year == 2024) |> pull(frac_urm)

# Long-term trend: 2010 vs 2024 first-year URM fraction
baruch_urm_2010 <- baruch_data |>
  filter(level == "first year undergrad", year == 2010) |>
  pull(frac_urm)

# ── California comparison (UCLA as reference) ─────────────────────────────────
ucla_data <- IPEDS |>
  filter(str_detect(institution_name, "Los Angeles") & is_calpublic)

ucla_fy_urm_2023 <- ucla_data |>
  filter(level == "first year undergrad", year == 2023) |>
  pull(frac_urm)

ucla_fy_urm_2024 <- ucla_data |>
  filter(level == "first year undergrad", year == 2024) |>
  pull(frac_urm)

ucla_urm_change <- ucla_fy_urm_2024 - ucla_fy_urm_2023

# ── National first-year URM average: 2023 vs 2024 ────────────────────────────
national_urm <- IPEDS |>
  filter(level == "first year undergrad", year %in% c(2023, 2024)) |>
  group_by(year) |>
  summarize(
    total_urm      = sum(total_urm,      na.rm = TRUE),
    total_enrolled = sum(total_enrolled, na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(frac_urm = total_urm / total_enrolled)

nat_urm_2023 <- national_urm |> filter(year == 2023) |> pull(frac_urm)
nat_urm_2024 <- national_urm |> filter(year == 2024) |> pull(frac_urm)
nat_urm_change <- nat_urm_2024 - nat_urm_2023
```

---

## Final Deliverable: Op-Ed

---

### One Year After SFFA: What the Numbers Tell Us About Diversity at Baruch

*By the President of CUNY Bernard M. Baruch College*

---

The Supreme Court's June 2023 ruling in *Students for Fair Admissions v. Harvard*
sent a shockwave through American higher education. At Baruch College, we spent
the months that followed asking a deceptively simple question: what, concretely,
would change? Now that the first class admitted under the post-SFFA regime has
enrolled, we have data. Some findings reassure me. Others demand our honest
attention.

**How we define diversity.** For purposes of this analysis, I use the fraction of
enrolled students who identify as Black, Hispanic, American Indian or Alaskan
Native, Native Hawaiian or Pacific Islander, or two or more races — a grouping
commonly understood as underrepresented minorities (URM) in higher education.
This metric is imperfect; New York City is a city without a racial majority, and
a single number cannot capture the full texture of a community. But it is the
measure most directly implicated by SFFA, and it allows a consistent comparison
across institutions and years.

**The entering class.** In Fall 2024, Baruch enrolled
`r comma(baruch_fy_2024)` first-year, first-time undergraduates. That is
`r ifelse(baruch_fy_2024 > baruch_fy_2023, "an increase", "a decrease")` from the
`r comma(baruch_fy_2023)` who enrolled in Fall 2023. Of our 2024 entering class,
`r percent(baruch_urm_frac_2024, accuracy = 0.1)` identified as URM, compared
with `r percent(baruch_urm_frac_2023, accuracy = 0.1)` the year before — a
shift of `r ifelse(baruch_urm_change >= 0, "+", "")`
`r percent(abs(baruch_urm_change), accuracy = 0.1)` percentage points.

**The full student body.** Zooming out to all enrolled undergraduates — not just
first-year students — the picture is steadier. In 2024, the URM share of
Baruch's undergraduate population was
`r percent(baruch_ug_urm_2024, accuracy = 0.1)`, compared with
`r percent(baruch_ug_urm_2023, accuracy = 0.1)` in 2023. Multi-year cohorts
absorb single-year shocks, which is precisely why watching both metrics matters:
the entering class is our leading indicator; the full population is our lagging
one.

**The long view.** In 2010, URM students made up
`r percent(baruch_urm_2010, accuracy = 0.1)` of Baruch's entering class.
That figure stands at `r percent(baruch_urm_frac_2024, accuracy = 0.1)` today.
The multi-decade arc of our diversity work is real and it did not begin or end
with a single Supreme Court decision. Baruch has always drawn strength from
being a gateway institution for first-generation students, immigrants, and
New Yorkers who are underrepresented in elite higher education. That mission
predates SFFA and it will outlast it.

**The California test.** Perhaps the most instructive comparison is with
California's public universities. Under Proposition 209, passed in 1996,
California's public institutions have been barred from race-conscious admissions
for nearly three decades. They had no policy to change when SFFA was handed down.
If demographic shifts at those schools after 2023 mirror shifts at schools that
*did* use race-conscious review, then the ruling's practical impact is modest; if
SFFA schools diverge sharply from California, that divergence is the signal.

At UCLA the URM share of first-year undergraduates
moved from `r percent(ucla_fy_urm_2023, accuracy = 0.1)` in 2023 to
`r percent(ucla_fy_urm_2024, accuracy = 0.1)` in 2024, a change of
`r ifelse(ucla_urm_change >= 0, "+", "")`
`r percent(abs(ucla_urm_change), accuracy = 0.1)` percentage points. Baruch's
corresponding change was
`r ifelse(baruch_urm_change >= 0, "+", "")`
`r percent(abs(baruch_urm_change), accuracy = 0.1)` percentage points.
Nationally, first-year URM enrollment shifted by
`r ifelse(nat_urm_change >= 0, "+", "")`
`r percent(abs(nat_urm_change), accuracy = 0.1)` percentage points between 2023
and 2024. Taken together, these figures suggest that one year is too short a
window for definitive conclusions. The data is noisy; demographic inertia is
real; and more admissions cycles will be required before the true signal emerges
from the noise.

**Looking ahead.** At Baruch, we are not waiting passively for the data to
clarify. We are expanding partnerships with New York City public high schools,
increasing need-based financial aid, and investing in transfer pathways from CUNY
community colleges. Race-neutral pathways to diversity exist but they are simply
harder and more expensive than race-conscious ones. We intend to use every one of
them.

The first year after SFFA is a data point, not a verdict. We will be watching
closely, and we will be transparent about what we find.

---

## AI Usage Statement

::: {.callout-note title="AI Usage Statement" collapse="true"}

Ai (chat gpt + qwen) was used to help contextualize the ask and provided code and syntax  editorial suggestions in completing this
project.


:::
