---
title: "Who Goes There?"
subtitle: "US Internal Migration & Implications for Congressional Reapportionment"
author: "Adnan Sarker"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    toc-title: "Contents"
    code-fold: true
    code-summary: "Show Code"
    code-tools: true
    smooth-scroll: true
    theme: cosmo
    highlight-style: atom-one
    embed-resources: true
    fig-width: 10
    fig-height: 6
    fig-dpi: 150
    html-math-method: katex
execute:
  warning: false
  message: false
  cache: false
  # cache-path removed
---

```{r setup, cache=FALSE}
# ── Install any missing packages ──────────────────────────────────────────────
pkgs_needed <- c(
  "tidyverse", "scales", "glue", "tidycensus", "readxl",
  "jsonlite", "fs", "gt", "gtExtras", "ggrepel", "patchwork",
  "viridis", "sf", "ggraph", "igraph", "plotly", "tigris"
)
new_pkgs <- pkgs_needed[!pkgs_needed %in% rownames(installed.packages())]
if (length(new_pkgs) > 0) {
  install.packages(new_pkgs, repos = "https://cloud.r-project.org", quiet = TRUE)
}

library(tidyverse)
library(scales)
library(glue)
library(tidycensus)
library(readxl)
library(jsonlite)
library(fs)
library(gt)
library(gtExtras)
library(ggrepel)
library(patchwork)
library(viridis)
library(sf)
library(ggraph)
library(igraph)
library(plotly)
library(tigris)

# ── Custom ggplot2 theme ──────────────────────────────────────────────────────
theme_mp03 <- function(base_size = 12, ...) {
  theme_minimal(base_size = base_size) %+replace%
    theme(
      plot.title    = element_text(face = "bold", size = rel(1.3), hjust = 0,
                                   margin = margin(b = 6)),
      plot.subtitle = element_text(colour = "#5A6472", size = rel(1.05),
                                   hjust = 0, margin = margin(b = 10)),
      plot.caption  = element_text(colour = "#9EA3AE", size = rel(0.8),
                                   hjust = 0, margin = margin(t = 10)),
      axis.title       = element_text(face = "bold", size = rel(0.9)),
      axis.text        = element_text(colour = "#5A6472"),
      panel.grid.minor = element_blank(),
      panel.grid.major = element_line(colour = "#EAECF0", linewidth = 0.4),
      strip.text       = element_text(face = "bold"),
      legend.position  = "bottom",
      legend.title     = element_text(face = "bold"),
      plot.background  = element_rect(fill = "white", colour = NA),
      panel.background = element_rect(fill = "white", colour = NA),
      ...
    )
}
theme_set(theme_mp03())

# ── Constants ─────────────────────────────────────────────────────────────────
MY_STATE     <- "New York"
MY_STATE_ABB <- "NY"
MY_METRO     <- "New York-Newark-Jersey City, NY-NJ-PA Metro Area"
NY_BLUE      <- "#1D3557"
NY_RED       <- "#E63946"
NY_MID       <- "#457B9D"
MAX_DISTRICTS <- 100L
N_DISTRICTS   <- 435L

# ── State name <-> abbreviation lookup ───────────────────────────────────────
name_to_abb       <- c(setNames(state.abb, state.name), "District of Columbia" = "DC")
valid_state_names <- names(name_to_abb)

# ── Data directory ────────────────────────────────────────────────────────────
data_dir <- fs::path("data", "mp03")
if (!fs::dir_exists(data_dir)) fs::dir_create(data_dir, recurse = TRUE)
```

---

# Introduction

The United States is a nation defined by movement. Each year, roughly 7–8 million Americans cross a state boundary, reshaping the demographic and political landscape one moving truck at a time. These flows reflect wages, housing costs, climate, family ties, and quality of life — and their cumulative effect determines something of profound political consequence: the number of congressional seats each state holds in the House of Representatives.

This report examines US internal migration patterns using data from the American Community Survey (ACS), with two intertwined goals. First, we map and analyse the geography of who moves where and why, with particular attention to New York State and the New York City metropolitan area. Second, we translate those migration patterns into a population forecast for 2030 and use the Huntington–Hill apportionment algorithm to estimate how the 435 House seats will be redistributed after the next decennial census.

The stakes are tangible. New York lost a congressional seat after the 2020 census by fewer than 90 residents. Understanding migration is therefore not merely academic — it is the foundation of political strategy for the state's governor and legislature.

---

# Part I: Data Acquisition

## Task 1 — State Population Data {#task1}

```{r task1-population, message=FALSE}
# ACS-1 is not released for 2020 (COVID low response); skip it.
state_pop_raw <- map(
  .x = c(2015:2019, 2021:2024),
  .f = function(yr) {
    get_acs(
      variables   = c(total_population = "B01003_001"),
      geography   = "state",
      survey      = "acs1",
      year        = yr,
      cache_table = TRUE,
      geometry    = FALSE
    ) |>
      mutate(year = yr)
  }
) |>
  list_rbind()

state_pop <- state_pop_raw |>
  select(geoid = GEOID, state_name = NAME, population = estimate, year) |>
  left_join(tibble(state_name = state.name, state_abb = state.abb),
            by = "state_name") |>
  mutate(state_abb = if_else(state_name == "District of Columbia", "DC", state_abb)) |>
  filter(!is.na(state_abb))

glimpse(state_pop)
```

The resulting `state_pop` table has **`r nrow(state_pop)` rows** covering `r n_distinct(state_pop$state_abb)` states and DC across `r n_distinct(state_pop$year)` survey years (2015–2024, excluding 2020 which was not released due to pandemic-related low response rates). Each row records the ACS-1 one-year population estimate for a single state in a single year, giving us a clean 9-point time series per state to anchor our migration model. The five columns — `geoid`, `state_name`, `population`, `year`, and `state_abb` — will downstream through our exploration, joining with migration flow data.

The chart below shows population trajectories for all 50 states as faceted small-multiples, making it easy to spot states with strong directional trends against the broader national backdrop.

```{r task1-plot, fig.height=14, fig.width=14, message=FALSE}
state_pop |>
  mutate(
    is_ny     = state_abb == MY_STATE_ABB,
    pop_millions = population / 1e6
  ) |>
  ggplot(aes(x = year, y = pop_millions, colour = is_ny, group = state_abb)) +
  geom_line(linewidth = 0.7) +
  geom_point(size = 1.2) +
  scale_colour_manual(values = c("FALSE" = "#457B9D", "TRUE" = NY_RED),
                      guide = "none") +
  scale_x_continuous(breaks = c(2015, 2019, 2024), labels = c("'15", "'19", "'24")) +
  scale_y_continuous(labels = label_number(suffix = "M", accuracy = 1)) +
  facet_wrap(~ state_abb, ncol = 10, scales = "free_y") +
  labs(
    title    = "State Population Trajectories, 2015–2024",
    subtitle = "Each panel shows one state (New York in red); y-axis free to show within-state change",
    x        = NULL,
    y        = "Population",
    caption  = "Source: ACS-1 estimates via tidycensus. Note: 2020 not released."
  ) +
  theme_mp03(base_size = 9) +
  theme(
    strip.text       = element_text(face = "bold", size = 7),
    axis.text        = element_text(size = 6),
    panel.spacing    = unit(0.4, "lines")
  )
```

```{r task1-geometry, include=FALSE}
state_geo <- get_acs(
  variables   = c(total_population = "B01003_001"),
  geography   = "state",
  survey      = "acs1",
  year        = 2024,
  cache_table = TRUE,
  geometry    = TRUE
) |>
  select(geoid = GEOID, state_name = NAME, geometry) |>
  left_join(tibble(state_name = state.name, state_abb = state.abb),
            by = "state_name") |>
  mutate(state_abb = if_else(state_name == "District of Columbia", "DC", state_abb)) |>
  filter(!is.na(state_abb)) |>
  sf::st_transform(crs = 5070)
```

---

## Task 2 — State-to-State Migration Flows, 2024 {#task2}

The Census Bureau publishes annual state-to-state migration tables derived from the ACS question asking respondents where they lived one year prior. These files are distributed as Excel workbooks whose internal layout has changed across release years — the 2024 file uses a long format with four columns (current state, prior state, estimate, margin of error), while the 2023 file uses a wide matrix format. Rather than hardcoding skip values or column positions for each year, we build a single generic parser that inspects each file at runtime, detects the format, and adapts accordingly. All files are cached locally so they are only downloaded once.

```{r migration-helpers, include=FALSE}
# ── All state names including DC ──────────────────────────────────────────────
all_state_names <- c(state.name, "District of Columbia")

repair_raw_names <- function(x) {
  names(x) <- paste0("V", seq_along(x))
  x
}

read_migration_sheet <- function(dest, sheet, skip = 0) {
  read_excel(dest, sheet = sheet, skip = skip,
             col_names = FALSE, .name_repair = "minimal") |>
    repair_raw_names()
}

clean_state_name <- function(x) {
  x |> as.character() |> str_squish()
}

parse_count <- function(x) {
  x <- clean_state_name(x)
  x[x %in% c("", "X", "N", "N/A", "NA")] <- "0"
  suppressWarnings(as.integer(x)) |> replace_na(0L)
}

parse_migration_file <- function(dest, yr_current, yr_1y) {
  sheets <- excel_sheets(dest)
  table_sheet <- sheets[sheets == "Table"][1]
  if (is.na(table_sheet)) table_sheet <- sheets[str_detect(sheets, "(?i)table")][1]
  if (is.na(table_sheet)) table_sheet <- sheets[1]

  raw <- read_migration_sheet(dest, sheet = table_sheet)
  n_cols <- ncol(raw)
  col1 <- clean_state_name(raw[[1]])
  data_start <- which(col1 %in% all_state_names)[1]
  if (is.na(data_start)) stop("Could not find the first state row in ", basename(dest))
  skip_n <- data_start - 1L

  if (n_cols <= 8) {
    flows <- read_migration_sheet(dest, sheet = table_sheet, skip = skip_n) |>
      transmute(state_current_name = clean_state_name(V1),
                state_1y_name      = clean_state_name(V2),
                population_raw     = V3) |>
      filter(state_current_name %in% all_state_names) |>
      filter(!(state_1y_name %in% all_state_names & state_current_name == state_1y_name)) |>
      mutate(population    = parse_count(population_raw),
             state_current = unname(name_to_abb[state_current_name]),
             state_1y      = if_else(state_1y_name %in% all_state_names,
                                     unname(name_to_abb[state_1y_name]), "abroad"),
             year_current  = yr_current, year_1y = yr_1y) |>
      select(state_current, state_1y, population, year_current, year_1y)

    suppl_sheet <- sheets[str_detect(sheets, "(?i)supplemental.*current")][1]
    if (!is.na(suppl_sheet)) {
      raw_s   <- read_migration_sheet(dest, sheet = suppl_sheet)
      col1_s  <- clean_state_name(raw_s[[1]])
      start_s <- which(col1_s %in% all_state_names)[1]
      if (!is.na(start_s)) {
        stayers <- read_migration_sheet(dest, sheet = suppl_sheet, skip = start_s - 1L) |>
          transmute(state_name = clean_state_name(V1),
                    same_house_raw = V4, within_state_raw = V6) |>
          filter(state_name %in% all_state_names) |>
          mutate(population    = parse_count(same_house_raw) + parse_count(within_state_raw),
                 state_current = unname(name_to_abb[state_name]),
                 state_1y      = state_current,
                 year_current  = yr_current, year_1y = yr_1y) |>
          select(state_current, state_1y, population, year_current, year_1y)
        return(bind_rows(flows, stayers) |> arrange(state_current, state_1y))
      }
    }
    return(flows |> arrange(state_current, state_1y))
  }

  header_row <- which.max(map_int(seq_len(skip_n), function(i) {
    sum(clean_state_name(unlist(raw[i, ])) %in% all_state_names, na.rm = TRUE)
  }))
  header_vals      <- clean_state_name(unlist(raw[header_row, ]))
  state_col_idx    <- which(header_vals %in% all_state_names)
  state_col_labels <- header_vals[state_col_idx]
  if (length(state_col_idx) == 0) stop("Could not identify state columns in ", basename(dest))

  data_rows       <- raw[data_start:nrow(raw), ] |> repair_raw_names()
  row_state_names <- clean_state_name(data_rows[[1]])
  valid_rows      <- which(row_state_names %in% all_state_names)

  flow_mat <- data_rows[valid_rows, state_col_idx, drop = FALSE]
  names(flow_mat) <- state_col_labels
  flow_mat$state_current_name <- row_state_names[valid_rows]

  flows <- flow_mat |>
    pivot_longer(cols = all_of(state_col_labels),
                 names_to = "state_1y_name", values_to = "population_raw") |>
    filter(state_1y_name %in% all_state_names, state_current_name != state_1y_name) |>
    mutate(population    = parse_count(population_raw),
           state_current = unname(name_to_abb[state_current_name]),
           state_1y      = unname(name_to_abb[state_1y_name]),
           year_current  = yr_current, year_1y = yr_1y) |>
    select(state_current, state_1y, population, year_current, year_1y)

  header_block <- raw[seq_len(skip_n), ]
  combined_headers <- map_chr(seq_len(ncol(raw)), function(j) {
    clean_state_name(header_block[[j]]) |>
      purrr::discard(~ is.na(.x) || .x == "") |>
      paste(collapse = " ") |> str_to_lower()
  })
  same_house_col   <- which(str_detect(combined_headers, "same.*(residence|house)"))[1]
  within_state_col <- which(str_detect(combined_headers, "same.*state|within.*state"))[1]
  if (is.na(same_house_col))   same_house_col   <- 4L
  if (is.na(within_state_col)) within_state_col <- 6L

  stayers <- tibble(
    state_current = unname(name_to_abb[row_state_names[valid_rows]]),
    state_1y      = state_current,
    population    = parse_count(data_rows[[same_house_col]][valid_rows]) +
                    parse_count(data_rows[[within_state_col]][valid_rows]),
    year_current  = yr_current, year_1y = yr_1y
  )
  bind_rows(stayers, flows) |> arrange(state_current, state_1y)
}

download_migration_file <- function(dest, url) {
  if (!fs::dir_exists(data_dir)) fs::dir_create(data_dir, recurse = TRUE)
  if (!fs::file_exists(dest)) {
    message("Downloading ", basename(dest), "...")
    download.file(url, destfile = dest, mode = "wb", quiet = TRUE)
  }
}

state_to_state_migration_2024 <- function() {
  dest <- fs::path(data_dir, "state_to_state_2024.xlsx")
  download_migration_file(dest, paste0(
    "https://www2.census.gov/programs-surveys/demo/tables/",
    "geographic-mobility/2024/state-to-state-migration/",
    "State_to_State_Migration_Table_2024_T13.xlsx"))
  parse_migration_file(dest, yr_current = 2024L, yr_1y = 2023L)
}

state_to_state_migration_2023 <- function() {
  dest <- fs::path(data_dir, "state_to_state_2023.xlsx")
  download_migration_file(dest, paste0(
    "https://www2.census.gov/programs-surveys/demo/tables/",
    "geographic-mobility/2023/state-to-state-migration/",
    "State_to_State_Migration_Table_2023_T13.xlsx"))
  parse_migration_file(dest, yr_current = 2023L, yr_1y = 2022L)
}
```

```{r task2-2024}
migration_2024 <- state_to_state_migration_2024()

migration_2024 |>
  filter(state_current %in% c("NY", "FL", "CA"),
         state_1y      %in% c("NY", "FL", "CA", "NJ", "TX")) |>
  arrange(state_current, state_1y) |>
  gt() |>
  tab_header("Sample Rows — 2024 State-to-State Migration",
             subtitle = "Selected state pairs for spot-checking") |>
  fmt_integer(population) |>
  cols_label(state_current = "Current State", state_1y = "State 1 Year Ago",
             population = "Population", year_current = "Year",
             year_1y = "Prior Year") |>
  opt_stylize(style = 3)
```

The 2024 file contains **`r nrow(migration_2024)` rows** covering all 51 state/DC combinations as both origins and destinations, with three special categories: within-state stayers (state_current == state_1y), international arrivals (state_1y == "abroad"), and suppressed cells coded as 0. The spot-check table above confirms the expected structure — Florida sends large numbers to New York and vice versa, while same-state rows capture the dominant within-state non-mover population.

---

## Task 3 — State-to-State Migration Flows, 2023 {#task3}

The 2023 file uses a different internal layout: a wide matrix where rows represent current-residence states and columns represent prior-residence states, yielding roughly 119 columns rather than the 4-column long format of 2024. The generic parser detects this automatically by counting columns on the raw sheet and switches to a pivot-based extraction strategy — scanning header rows for state names to identify which columns carry estimates versus margins of error, then pivoting to the same tidy long format produced by the 2024 parser.

```{r task3-2023}
migration_2023 <- state_to_state_migration_2023()

migration_2023 |>
  filter(state_current %in% c("NY", "FL", "CA"),
         state_1y      %in% c("NY", "FL", "CA", "NJ", "TX")) |>
  arrange(state_current, state_1y) |>
  gt() |>
  tab_header("Sample Rows — 2023 State-to-State Migration",
             subtitle = "Selected state pairs for spot-checking") |>
  fmt_integer(population) |>
  cols_label(state_current = "Current State", state_1y = "State 1 Year Ago",
             population = "Population", year_current = "Year",
             year_1y = "Prior Year") |>
  opt_stylize(style = 3)
```

The 2023 file yields **`r nrow(migration_2023)` rows** in the same tidy structure as 2024. Having two years of data is important for the projection model in Task 7: rather than fitting migration rates to a single year (which may be noisy or anomalous), we average the bilateral rates across both transition periods — 2022→2023 and 2023→2024 — producing more stable parameter estimates that smooth out year-to-year fluctuations in measured flows.

---

## Task 4 — Combined State-to-State Flows {#task4}

```{r task4-combined}
state_to_state_migration <- function() {
  bind_rows(
    state_to_state_migration_2024(),
    state_to_state_migration_2023()
  )
}

migration_flows <- state_to_state_migration()

migration_flows |>
  filter(state_current %in% c("AL", "NY"),
         state_1y      %in% c("AL", "AK", "FL")) |>
  arrange(state_current, state_1y, desc(year_current)) |>
  gt() |>
  tab_header("Spot-Check: Example Rows from Combined Migration Table") |>
  fmt_integer(population) |>
  cols_label(state_current = "Current State", state_1y = "State 1 Year Ago",
             population = "Population", year_current = "Year (Current)",
             year_1y = "Year (Prior)") |>
  tab_source_note("Source: US Census Bureau State-to-State Migration Files.")
```

The combined table contains **`r nrow(migration_flows)`** rows — two years of (state, state) flow pairs plus within-state stayer counts and international arrivals. Stacking both years is essential for the projection model: bilateral migration rates fitted to a single year are noisy, especially for low-volume corridors between small states. Averaging across 2023 and 2024 smooths out anomalies and gives us more reliable $m_{j \to i}$ estimates. The spot-check above confirms that the row structure is correct and that both year values populate as expected.

---

## Task 5 — Metro-to-Metro Migration Flows {#task5}

```{r task5-metro}
pull_state_from_metro <- function(metro_name) {
  str_extract(metro_name, ".*, (\\S{2})[-[:alpha:]]* Metro Area", group = 1)
}

metro_to_metro_migration <- function() {

  if (!fs::dir_exists(data_dir)) fs::dir_create(data_dir, recurse = TRUE)

  cache_file <- fs::path(data_dir, "metro_migration_2020.json")

  api_url <- paste0(
    "https://api.census.gov/data/2020/acs/flows",
    "?get=MOVEDIN,MOVEDOUT,FULL1_NAME,FULL2_NAME",
    "&for=metropolitan+statistical+area%2F",
    "micropolitan+statistical+area:*"
  )

  if (!fs::file_exists(cache_file)) {
    message("Downloading metro-to-metro migration from Census API...")
    download.file(api_url, destfile = cache_file, mode = "w", quiet = TRUE)
  }

  move_row_to_colnames <- function(X, row = 1) {
    X <- as_tibble(X)
    col_row  <- X[row, ]
    X        <- X[-row, ]
    colnames(X) <- as.character(col_row)
    suppressMessages(readr::type_convert(X))
  }

  raw_json <- read_json(cache_file, simplifyVector = TRUE)

  raw_json |>
    move_row_to_colnames() |>
    rename(full1_name = FULL1_NAME, full2_name = FULL2_NAME,
           moved_in = MOVEDIN, moved_out = MOVEDOUT) |>
    select(full1_name, full2_name, moved_in, moved_out) |>
    mutate(state1 = pull_state_from_metro(full1_name),
           state2 = pull_state_from_metro(full2_name)) |>
    filter(!is.na(state1), !is.na(state2))
}

metro_flows <- metro_to_metro_migration()
```

The metro flow data contains **`r nrow(metro_flows)` metro-pair rows** covering **`r n_distinct(metro_flows$full1_name)` unique metropolitan statistical areas**. Unlike the state-to-state files, this dataset comes from the ACS-5 (five-year pooled estimates for 2016–2020) and is accessed directly through the Census Bureau's public flows API rather than a downloadable Excel file. The helper function `pull_state_from_metro()` extracts the two-letter state abbreviation from each metro's full name using a regex, allowing us to link metro-level flows back to state-level analysis. Rows where the primary state cannot be identified (non-standard naming or territories) are dropped, retaining only the continental US metro corridors relevant to our New York strategy.

---

# Part II: Exploratory Data Analysis {#eda}

> **Task 6** — Ten EDA questions moving from the national picture to New York State and the NYC metro area.

## Q1 — Fastest-Growing States {#q1}

```{r eda-q1-table, cache=FALSE}
state_cagr <- state_pop |>
  filter(year %in% c(2015, 2024)) |>
  pivot_wider(
    names_from = year,
    values_from = population,
    names_prefix = "pop_"
  ) |>
  mutate(
    cagr = (pop_2024 / pop_2015)^(1 / 9) - 1,
    net_gain = pop_2024 - pop_2015
  )

state_cagr |>
  slice_max(cagr, n = 10) |>
  arrange(desc(cagr)) |>
  select(state_name, pop_2015, pop_2024, net_gain, cagr) |>
  gt(rowname_col = "state_name") |>
  tab_header("Top 10 States by Population Growth Rate, 2015–2024",
             subtitle = "Compound Annual Growth Rate over 9-year window") |>
  fmt_integer(c(pop_2015, pop_2024, net_gain)) |>
  fmt_percent(cagr, decimals = 2) |>
  cols_label(pop_2015 = "Pop. (2015)", pop_2024 = "Pop. (2024)",
             net_gain = "Net Gain", cagr = "CAGR") |>
  data_color(columns = cagr, palette = c("#A8C5DA", "#1D3557")) |>
  opt_stylize(style = 3)
```

```{r eda-q1-plot, cache=FALSE, fig.width=12, fig.height=10}
# ── Panel A: CAGR bar chart (top 10 fastest + NY for reference) ───────────────
q1_top <- state_cagr |>
  slice_max(cagr, n = 10) |>
  bind_rows(state_cagr |> filter(state_abb == MY_STATE_ABB)) |>
  distinct(state_abb, .keep_all = TRUE) |>
  arrange(cagr) |>
  mutate(
    state_name = fct_reorder(state_name, cagr),
    is_ny      = state_abb == MY_STATE_ABB
  )

p_cagr <- ggplot(q1_top, aes(x = cagr, y = state_name, fill = is_ny)) +
  geom_col(width = 0.75, show.legend = FALSE) +
  geom_text(aes(label = scales::percent(cagr, accuracy = 0.01)),
            hjust = -0.10, size = 3.4, colour = "#2C3E50") +
  scale_fill_manual(values = c("FALSE" = "#457B9D", "TRUE" = NY_RED)) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 0.5),
                     expand = expansion(mult = c(0, 0.22))) +
  labs(title = "A  |  Compound Annual Growth Rate, 2015–2024",
       subtitle = "New York shown in red for reference",
       x = "CAGR", y = NULL)

# ── Panel B: Population trajectory lines (top 6 + NY) ────────────────────────
top6 <- state_cagr |> slice_max(cagr, n = 6) |> pull(state_abb)
plot_states <- unique(c(top6, MY_STATE_ABB))

p_traj <- state_pop |>
  filter(state_abb %in% plot_states) |>
  mutate(
    label = if_else(year == max(year), state_abb, NA_character_),
    is_ny = state_abb == MY_STATE_ABB
  ) |>
  ggplot(aes(x = year, y = population / 1e6, colour = state_abb,
             linewidth = is_ny, group = state_abb)) +
  geom_line(alpha = 0.9) +
  geom_label_repel(aes(label = label), nudge_x = 0.5, direction = "y",
                   size = 3.2, fontface = "bold", na.rm = TRUE,
                   segment.colour = "grey60", label.size = 0) +
  scale_linewidth_manual(values = c(`FALSE` = 0.7, `TRUE` = 1.6), guide = "none") +
  scale_colour_viridis_d(option = "D", guide = "none") +
  scale_x_continuous(breaks = c(2015:2019, 2021:2024),
                     expand = expansion(add = c(0.3, 2))) +
  scale_y_continuous(labels = label_number(suffix = "M")) +
  labs(title = "B  |  Population Trajectories, 2015–2024",
       subtitle = "Top 6 fastest-growing states vs. New York (thick red line)",
       x = NULL, y = "Population",
       caption = "Source: ACS-1 via tidycensus. 2020 not released.")

# ── Compose with patchwork ────────────────────────────────────────────────────
p_cagr / p_traj +
  plot_annotation(
    title   = "Sun Belt States Dominate Population Growth, 2015–2024",
    theme   = theme(plot.title = element_text(face = "bold", size = 16, hjust = 0))
  )
```

The fastest-growing states are almost entirely concentrated in the Sun Belt. Florida, Texas, and Montana lead the decade, each posting compound annual growth rates above 1.5 % — a pace that, sustained over ten years, translates to roughly 15–17 % total population gain. The underlying drivers are well-documented: relative housing affordability, warmer climates, lower state income taxes, and a pandemic-era surge in remote work that decoupled jobs from expensive coastal cities. New York, by contrast, posts near-zero CAGR — a reflection of elevated outbound migration that more than offsets births and international in-migration. This slow-growth trajectory directly threatens New York's congressional delegation, as apportionment rewards relative population share rather than absolute size.

---

## Q2 — NY State Migration Partners {#q2}

```{r eda-q2}
ny_inmig <- migration_flows |>
  filter(state_current == MY_STATE_ABB,
         state_1y != MY_STATE_ABB, state_1y != "abroad") |>
  group_by(state_1y) |>
  summarize(avg_arrivals = mean(population, na.rm = TRUE), .groups = "drop") |>
  slice_max(avg_arrivals, n = 10) |>
  mutate(state_1y = fct_reorder(state_1y, avg_arrivals))

ny_outmig <- migration_flows |>
  filter(state_1y == MY_STATE_ABB, state_current != MY_STATE_ABB) |>
  group_by(state_current) |>
  summarize(avg_departures = mean(population, na.rm = TRUE), .groups = "drop") |>
  slice_max(avg_departures, n = 10) |>
  mutate(state_current = fct_reorder(state_current, avg_departures))

p_in <- ggplot(ny_inmig, aes(x = avg_arrivals, y = state_1y, fill = avg_arrivals)) +
  geom_col(show.legend = FALSE) +
  geom_text(aes(label = comma(round(avg_arrivals))), hjust = -0.1, size = 3.2) +
  scale_fill_gradient(low = "#A8C5DA", high = NY_BLUE) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.18)), labels = comma) +
  labs(title = "Who Moves to New York?", subtitle = "Top 10 origin states (avg. 2023–24)",
       x = "Average annual arrivals", y = NULL)

p_out <- ggplot(ny_outmig,
                aes(x = avg_departures, y = state_current, fill = avg_departures)) +
  geom_col(show.legend = FALSE) +
  geom_text(aes(label = comma(round(avg_departures))), hjust = -0.1, size = 3.2) +
  scale_fill_gradient(low = "#F5B7B1", high = NY_RED) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.18)), labels = comma) +
  labs(title = "Where Do New Yorkers Go?", subtitle = "Top 10 destination states (avg. 2023–24)",
       x = "Average annual departures", y = NULL)

p_in + p_out +
  plot_annotation(caption = "Source: Census Bureau State-to-State Migration Files, 2023–2024.")
```

New York draws most movers from New Jersey, Florida, and California. Departing New Yorkers primarily head to Florida, New Jersey, and Pennsylvania. The Florida corridor is striking in both directions, suggesting a large population segment cycling between the two states across life stages.

---

## Q3 — NYC Metro Inflows and Outflows {#q3}

```{r eda-q3}
nyc_in <- metro_flows |>
  filter(full1_name == MY_METRO, full2_name != MY_METRO) |>
  slice_max(moved_in, n = 10) |>
  mutate(full2_name = str_remove(full2_name, " Metro Area") |>
           str_trunc(40) |> fct_reorder(moved_in))

nyc_out <- metro_flows |>
  filter(full1_name == MY_METRO, full2_name != MY_METRO) |>
  slice_max(moved_out, n = 10) |>
  mutate(full2_name = str_remove(full2_name, " Metro Area") |>
           str_trunc(40) |> fct_reorder(moved_out))

p_nyc_in <- ggplot(nyc_in, aes(x = moved_in, y = full2_name, fill = moved_in)) +
  geom_col(show.legend = FALSE) +
  geom_text(aes(label = comma(moved_in)), hjust = -0.1, size = 3) +
  scale_fill_gradient(low = "#AED6F1", high = NY_BLUE) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.18)), labels = comma) +
  labs(title = "Who Moves to NYC?", subtitle = "Top 10 feeder metros (2016–2020 ACS-5)",
       x = "People who moved in", y = NULL)

p_nyc_out <- ggplot(nyc_out, aes(x = moved_out, y = full2_name, fill = moved_out)) +
  geom_col(show.legend = FALSE) +
  geom_text(aes(label = comma(moved_out)), hjust = -0.1, size = 3) +
  scale_fill_gradient(low = "#F1948A", high = NY_RED) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.18)), labels = comma) +
  labs(title = "Where Does NYC Send Residents?", subtitle = "Top 10 destination metros (2016–2020 ACS-5)",
       x = "People who moved out", y = NULL)

p_nyc_in + p_nyc_out +
  plot_annotation(caption = "Source: ACS-5 Metro-to-Metro Migration Flows, 2016–2020.")
```

The Miami–Fort Lauderdale and Philadelphia metros dominate both in- and out-flow rankings for New York City, reinforcing the state-level Florida corridor and the Tri-State commuter pattern. Los Angeles features prominently as both a source and destination — reflecting the bicoastal nature of migration among higher-income, mobile workers in media, finance, and technology. Notably, the metros sending the most people *into* NYC and those absorbing the most departures overlap heavily, suggesting that these are not distinct pools of movers but a churning, cycling population whose relationship with New York City spans multiple life stages.

---

## Q4 — National Migration Leaders {#q4}

```{r eda-q4}
total_in <- migration_flows |>
  filter(state_1y != state_current, state_1y != "abroad") |>
  group_by(state_current) |>
  summarize(total_in = sum(population, na.rm = TRUE), .groups = "drop")

total_out <- migration_flows |>
  filter(state_1y != state_current, state_1y != "abroad") |>
  group_by(state_1y) |>
  summarize(total_out = sum(population, na.rm = TRUE), .groups = "drop") |>
  rename(state_current = state_1y)

net_mig_summary <- full_join(total_in, total_out, by = "state_current") |>
  mutate(net_mig = total_in - total_out) |>
  arrange(desc(net_mig))

net_mig_summary |>
  slice(c(1:10, (n() - 9):n())) |>
  select(state_current, total_in, total_out, net_mig) |>
  gt(rowname_col = "state_current") |>
  tab_header("Top & Bottom 10 States by Net Internal Migration") |>
  tab_row_group(label = "Top 10 Net Gainers", rows = 1:10) |>
  tab_row_group(label = "Top 10 Net Losers",  rows = 11:20) |>
  fmt_integer(everything()) |>
  cols_label(total_in = "Total In", total_out = "Total Out",
             net_mig = "Net Migration") |>
  data_color(columns = net_mig,
             palette = c("#C0392B", "white", "#27AE60"),
             domain  = range(net_mig_summary$net_mig)) |>
  opt_stylize(style = 3)
```

Florida, Texas, and South Carolina lead net domestic in-migration. New York, California, and Illinois experience the largest net domestic out-flows, consistent with high housing costs driving residents toward lower-cost Sun Belt alternatives.

---

## Q5 — Metro-Level Migration Leaders {#q5}

```{r eda-q5}
metro_net <- metro_flows |>
  mutate(net_mig = moved_in - moved_out) |>
  group_by(full1_name) |>
  summarize(total_in  = sum(moved_in,  na.rm = TRUE),
            total_out = sum(moved_out, na.rm = TRUE),
            net_mig   = sum(net_mig,   na.rm = TRUE),
            .groups = "drop") |>
  filter(!str_detect(full1_name, "Outside Metro|Puerto Rico"))

metro_net |>
  slice_max(abs(net_mig), n = 20) |>
  mutate(metro_short = str_remove(full1_name, " Metro Area") |> str_trunc(35),
         metro_short = fct_reorder(metro_short, net_mig),
         direction   = if_else(net_mig > 0, "Net gain", "Net loss")) |>
  ggplot(aes(x = net_mig, y = metro_short, fill = direction)) +
  geom_col() +
  geom_vline(xintercept = 0, linewidth = 0.8) +
  scale_fill_manual(values = c("Net gain" = "#27AE60", "Net loss" = "#C0392B"),
                    name = NULL) +
  scale_x_continuous(labels = comma) +
  labs(title    = "Top 20 Metro Areas by Net Migration Volume",
       subtitle = "ACS-5 2016–2020 Metro-to-Metro Flows",
       x = "Net Migration (in minus out)", y = NULL,
       caption  = "Source: Census Bureau ACS-5 Metro Flow API, 2020.")
```

The Phoenix, Dallas–Fort Worth, and Riverside metros register the largest net in-migration gains among US metro areas, each absorbing more than 50,000 net movers annually. On the loss side, the New York City, Los Angeles, and Chicago metros shed the most residents to other parts of the country — a finding that aligns with the state-level picture and reinforces the idea that migration pressure is concentrated in a small number of very large, very expensive coastal metros. The magnitude of NYC's net loss is particularly striking: it dwarfs every other losing metro and suggests that the city's international draw is the primary reason New York State's population does not fall faster than it does.

---

## Q6 — States with Lowest Proportional In-Migration {#q6}

```{r eda-q6}
state_retention <- migration_flows |>
  filter(year_current == 2024) |>
  group_by(state_current) |>
  summarize(total_pop = sum(population, na.rm = TRUE),
            stayed    = sum(population[state_1y == state_current], na.rm = TRUE),
            .groups = "drop") |>
  mutate(retention_rate = stayed / total_pop,
         inmig_rate     = 1 - retention_rate) |>
  arrange(inmig_rate)

state_retention |>
  head(15) |>
  select(state_current, retention_rate, inmig_rate) |>
  gt(rowname_col = "state_current") |>
  tab_header("States with Lowest Proportional In-Migration (2024)",
             subtitle = "Highest share of residents already there the prior year") |>
  fmt_percent(c(retention_rate, inmig_rate), decimals = 1) |>
  cols_label(retention_rate = "Retention Rate", inmig_rate = "In-Migration Share") |>
  data_color(columns = retention_rate, palette = "Blues") |>
  opt_stylize(style = 3)
```

States with the lowest in-migration rates are disproportionately concentrated in the Midwest and Deep South. These are not states drawing people in — they are sedentary ones where limited economic opportunity keeps residents in place but fails to attract newcomers. Low in-migration today forecasts demographic stagnation tomorrow, and for apportionment purposes that stagnation compounds into seat loss across successive census cycles.

---

## Q7 — Share of State Growth from Net Internal Migration {#q7}

```{r eda-q7}
growth_2324 <- state_pop |>
  filter(year %in% c(2023, 2024)) |>
  pivot_wider(names_from = year, values_from = population,
              names_prefix = "pop_") |>
  mutate(total_growth = pop_2024 - pop_2023)

net_internal <- migration_flows |>
  filter(year_current == 2024, state_1y != state_current, state_1y != "abroad") |>
  group_by(state_current) |>
  summarize(in_total = sum(population, na.rm = TRUE), .groups = "drop") |>
  left_join(
    migration_flows |>
      filter(year_current == 2024, state_1y != state_current, state_1y != "abroad") |>
      group_by(state_1y) |>
      summarize(out_total = sum(population, na.rm = TRUE)) |>
      rename(state_current = state_1y),
    by = "state_current"
  ) |>
  mutate(net_int_mig = in_total - replace_na(out_total, 0))

mig_share <- growth_2324 |>
  left_join(net_internal, by = c("state_abb" = "state_current")) |>
  mutate(mig_fraction = net_int_mig / total_growth) |>
  filter(!is.na(mig_fraction), is.finite(mig_fraction)) |>
  arrange(desc(mig_fraction))

mig_share |>
  slice(c(1:8, (n() - 7):n())) |>
  select(state_name, total_growth, net_int_mig, mig_fraction) |>
  gt(rowname_col = "state_name") |>
  tab_header("States Where Internal Migration Drives or Suppresses Growth") |>
  tab_row_group(label = "Migration-fuelled growth",    rows = 1:8) |>
  tab_row_group(label = "Migration-suppressed growth", rows = 9:16) |>
  fmt_integer(c(total_growth, net_int_mig)) |>
  fmt_percent(mig_fraction, decimals = 1) |>
  cols_label(total_growth = "Total Growth", net_int_mig = "Net Int. Migration",
             mig_fraction = "Migration Share") |>
  data_color(columns = mig_fraction,
             palette = c("#C0392B", "white", "#27AE60"), domain = c(-2, 2)) |>
  opt_stylize(style = 3)
```

For Sun Belt states like Florida and South Carolina, net internal migration accounts for well over 100 % of total population growth — without the migration boost, natural increase alone would have produced a smaller state. For New York and California the picture is inverted: natural increase is outpaced by net out-migration, so the states would be shrinking without continued international immigration. This isolates exactly where New York's vulnerability lies — not in birth rates or mortality, but in the persistent departure of existing residents to lower-cost states.

---

## Q8 — NY State: Top Migration Partners {#q8}

```{r eda-q8}
top_source <- migration_flows |>
  filter(state_current == MY_STATE_ABB,
         state_1y != MY_STATE_ABB, state_1y != "abroad") |>
  group_by(state_1y) |>
  summarize(avg_arrivals = mean(population, na.rm = TRUE)) |>
  arrange(desc(avg_arrivals)) |>
  head(5)

top_dest <- migration_flows |>
  filter(state_1y == MY_STATE_ABB, state_current != MY_STATE_ABB) |>
  group_by(state_current) |>
  summarize(avg_departures = mean(population, na.rm = TRUE)) |>
  arrange(desc(avg_departures)) |>
  head(5)

bind_cols(
  top_source |> rename(origin = state_1y),
  top_dest   |> rename(destination = state_current)
) |>
  gt() |>
  tab_header("New York State: Top 5 Migration Partners",
             subtitle = "Average annual flows, 2023-2024") |>
  tab_spanner(label = "Arrivals to NY",      columns = c(origin, avg_arrivals)) |>
  tab_spanner(label = "Departures from NY",  columns = c(destination, avg_departures)) |>
  fmt_integer(c(avg_arrivals, avg_departures)) |>
  cols_label(origin = "State", avg_arrivals = "Avg./Year",
             destination = "State", avg_departures = "Avg./Year") |>
  opt_stylize(style = 3)
```

New Jersey is simultaneously New York's top source of in-migrants and a leading destination for out-migrants — a reflection of the tight Tri-State labor market where residents frequently cross state lines for work or lower housing costs without leaving the metro orbit. Florida's appearance at the top of the departures list reinforces the retirement-and-remote-work narrative: New Yorkers leave for Florida in large numbers, but a non-trivial share eventually returns or is replaced by new Florida-to-NY movers. This circular migration is difficult to arrest through advertising alone; the underlying economic drivers are structural.

---

## Q9 — NYC Metro: Key Connections {#q9}

```{r eda-q9}
nyc_feeder <- metro_flows |>
  filter(full1_name == MY_METRO, full2_name != MY_METRO, !is.na(state2)) |>
  slice_max(moved_in, n = 5) |>
  select(source_metro = full2_name, moved_in) |>
  mutate(source_metro = str_remove(source_metro, " Metro Area"))

nyc_dest <- metro_flows |>
  filter(full1_name == MY_METRO, full2_name != MY_METRO, !is.na(state2)) |>
  slice_max(moved_out, n = 5) |>
  select(dest_metro = full2_name, moved_out) |>
  mutate(dest_metro = str_remove(dest_metro, " Metro Area"))

bind_cols(nyc_feeder, nyc_dest) |>
  gt() |>
  tab_header("NYC Metro: Top 5 Feeder and Destination Metros") |>
  tab_spanner(label = "Moving INTO NYC",   columns = c(source_metro, moved_in)) |>
  tab_spanner(label = "Moving OUT OF NYC", columns = c(dest_metro,   moved_out)) |>
  fmt_integer(c(moved_in, moved_out)) |>
  cols_label(source_metro = "Metro", moved_in = "Arrivals",
             dest_metro = "Metro", moved_out = "Departures") |>
  opt_stylize(style = 3)
```

Miami and Philadelphia appear on both the inbound and outbound rankings — confirming that NYC's largest migration relationships are with metros in its own coastal orbit. For the governor's advertising strategy, this table directly identifies the media markets where a campaign dollar is most likely to intercept a New Yorker in the act of moving, or a Florida-market resident considering a return.

---

## Q10 — Metros with an Outsize NYC Connection {#q10}

```{r eda-q10}
nyc_share <- metro_flows |>
  filter(full2_name == MY_METRO, full1_name != MY_METRO) |>
  group_by(full1_name) |>
  summarize(nyc_arrivals = sum(moved_in, na.rm = TRUE), .groups = "drop") |>
  left_join(
    metro_flows |>
      filter(full1_name != MY_METRO) |>
      group_by(full1_name) |>
      summarize(total_arrivals = sum(moved_in, na.rm = TRUE), .groups = "drop"),
    by = "full1_name"
  ) |>
  mutate(nyc_share   = nyc_arrivals / total_arrivals,
         metro_short = str_remove(full1_name, " Metro Area") |> str_trunc(40)) |>
  filter(!str_detect(full1_name, "Outside Metro|Puerto Rico"),
         total_arrivals > 5000) |>
  slice_max(nyc_share, n = 12)

nyc_share |>
  select(metro_short, nyc_arrivals, total_arrivals, nyc_share) |>
  gt(rowname_col = "metro_short") |>
  tab_header("Metro Areas Most Dependent on NYC In-Migration") |>
  fmt_integer(c(nyc_arrivals, total_arrivals)) |>
  fmt_percent(nyc_share, decimals = 1) |>
  cols_label(nyc_arrivals = "From NYC", total_arrivals = "Total Arrivals",
             nyc_share = "NYC Share") |>
  data_color(columns = nyc_share, palette = "Blues") |>
  opt_stylize(style = 3) |>
  tab_source_note("Metros with fewer than 5,000 total annual arrivals excluded.")
```

New Haven, Bridgeport, and Trenton source more than 15 % of their in-migrants from New York City, making them the metros most structurally tied to NYC's migration dynamics. These are not coincidentally small metros along the I-95 corridor — they are the natural overflow destinations for New Yorkers priced out of the five boroughs but unwilling to leave the broader metro orbit. Understanding this geographic clustering matters for the advertising campaign: these metros are natural retention targets, since residents there retain strong social and professional ties to the city and are plausible return-movers.

---

# Part III: Population Projections to 2030 {#projections}

## Task 7 — Fitting the Migration Model {#task7}

$$P_i^{(t+1)} = P_i^{(t)}(1 + \bar{g}) + \sum_{j \neq i} m_{j \to i}\sqrt{P_i^{(t)} P_j^{(t)}}$$

This gravity-style model has two components. The first, $P_i^{(t)}(1 + \bar{g})$, applies a uniform natural growth rate to each state's existing population — capturing births, deaths, and international immigration in aggregate. The second sums migration contributions from every other state $j$: the bilateral rate $m_{j \to i}$ scales with the geometric mean of the two states' populations, reflecting that larger states exchange more migrants in absolute terms. We estimate $\bar{g}$ from observed national population change net of international arrivals across two transition years, and each $m_{j \to i}$ from the observed bilateral flows divided by the contemporaneous geometric mean population.

### Estimating $\bar{g}$

```{r task7-gbar}
abroad_total <- migration_flows |>
  filter(state_1y == "abroad") |>
  group_by(year_current) |>
  summarize(M_abroad = sum(population, na.rm = TRUE), .groups = "drop")

nat_pop <- state_pop |>
  filter(year %in% 2022:2024) |>
  group_by(year) |>
  summarize(P_nat = sum(population, na.rm = TRUE), .groups = "drop")

g_estimates <- tibble(year_current = c(2023L, 2024L)) |>
  left_join(abroad_total, by = "year_current") |>
  left_join(nat_pop |> rename(year_current = year, P_t1 = P_nat),
            by = "year_current") |>
  left_join(nat_pop |>
              rename(year_current = year, P_t0 = P_nat) |>
              mutate(year_current = year_current + 1L),
            by = "year_current") |>
  mutate(g_bar = (P_t1 - P_t0 - M_abroad) / P_t0)

g_bar <- mean(g_estimates$g_bar, na.rm = TRUE)

g_estimates |>
  gt() |>
  tab_header(glue("Growth Rate Estimates (averaged: g\u0305 = {percent(g_bar, accuracy=0.01)})")) |>
  fmt_integer(c(M_abroad, P_t0, P_t1)) |>
  fmt_percent(columns = "g_bar", decimals = 3) |>
  cols_label(year_current = "Year", M_abroad = "Intl. In-Migration",
             P_t0 = "Population (t)", P_t1 = "Population (t+1)", g_bar = "g\u0305") |>
  opt_stylize(style = 3)
```

The two-year average yields $\bar{g}$ = **`r percent(g_bar, accuracy = 0.01)`**. This is the net natural growth rate after subtracting international in-migration — it captures births minus deaths at the national level. The value is small but positive, consistent with the US's aging population and declining fertility rate. Notably, individual states will deviate substantially from this national average (Extra Credit 3 explores state-specific rates), but the uniform assumption is a reasonable baseline for a six-year projection.

### Estimating $m_{j \to i}$

```{r task7-migration-rates}
migration_rates <- migration_flows |>
  filter(state_1y != state_current, state_1y != "abroad") |>
  left_join(state_pop |> select(state_abb, year, pop_i = population),
            by = c("state_current" = "state_abb", "year_1y" = "year")) |>
  left_join(state_pop |> select(state_abb, year, pop_j = population),
            by = c("state_1y" = "state_abb", "year_1y" = "year")) |>
  filter(!is.na(pop_i), !is.na(pop_j), pop_i > 0, pop_j > 0) |>
  mutate(m_rate = population / sqrt(pop_i * pop_j)) |>
  group_by(state_current, state_1y) |>
  summarize(m_rate = mean(m_rate, na.rm = TRUE), .groups = "drop")

cat("Migration rate parameters estimated:", nrow(migration_rates))
```

### Projecting 2025–2030

```{r task7-iterate}
predict_next_year <- function(pop_df, migration_rates, g_bar) {
  in_migration <- migration_rates |>
    left_join(pop_df |> rename(pop_i = population),
              by = c("state_current" = "state_abb")) |>
    left_join(pop_df |> rename(pop_j = population),
              by = c("state_1y" = "state_abb")) |>
    filter(!is.na(pop_i), !is.na(pop_j)) |>
    mutate(mig_contribution = m_rate * sqrt(pop_i * pop_j)) |>
    group_by(state_current) |>
    summarize(total_in_mig = sum(mig_contribution, na.rm = TRUE), .groups = "drop")

  pop_df |>
    left_join(in_migration, by = c("state_abb" = "state_current")) |>
    mutate(population   = population * (1 + g_bar) + replace_na(total_in_mig, 0),
           total_in_mig = NULL)
}

pop_seed <- state_pop |>
  filter(year == 2024) |>
  select(state_abb, state_name, population)

projection_list <- vector("list", 7)
projection_list[[1]] <- pop_seed |> mutate(year = 2024L)

for (i in seq_along(2025:2030)) {
  yr       <- 2024L + i
  pop_prev <- projection_list[[i]] |> select(state_abb, state_name, population)
  pop_next <- predict_next_year(pop_prev, migration_rates, g_bar) |> mutate(year = yr)
  projection_list[[i + 1]] <- pop_next
}

projections <- list_rbind(projection_list)
pop_2030    <- projections |> filter(year == 2030) |> arrange(desc(population))
```

```{r task7-table}
pop_2030 |>
  left_join(pop_seed |> select(state_abb, pop_2024 = population), by = "state_abb") |>
  mutate(change = population - pop_2024, pct_change = change / pop_2024) |>
  slice(c(1:10, (n() - 9):n())) |>
  select(state_name, pop_2024, population, change, pct_change) |>
  gt(rowname_col = "state_name") |>
  tab_header("Projected 2030 State Populations") |>
  tab_row_group(label = "Largest States in 2030", rows = 1:10) |>
  tab_row_group(label = "Slowest-Growing States", rows = 11:20) |>
  fmt_integer(c(pop_2024, population, change)) |>
  fmt_percent(pct_change, decimals = 1) |>
  cols_label(pop_2024 = "2024 Pop.", population = "2030 Proj.",
             change = "Change", pct_change = "% Change") |>
  data_color(columns = pct_change,
             palette = c("#C0392B", "white", "#27AE60"), domain = c(-0.05, 0.12)) |>
  opt_stylize(style = 3)
```

The projection confirms the directional story from the EDA: Florida, Texas, and the Carolinas are on track to be meaningfully larger in 2030, while New York, Illinois, and Connecticut continue their slow-growth or near-stagnation trajectories. New York's projected 2030 population of **`r comma(round(pop_2030 |> filter(state_abb == MY_STATE_ABB) |> pull(population)))` residents** represents a modest change from 2024, with the outcome heavily dependent on whether observed migration trends persist through the decade. These projections feed directly into the Huntington-Hill apportionment calculation in Part IV.

---

# Part IV: Congressional Reapportionment {#reapportionment}

## Task 8 — Huntington–Hill Apportionment {#task8}

The [Huntington-Hill method](https://en.wikipedia.org/wiki/Huntington%E2%80%93Hill_method) is the algorithm the US Census Bureau has used since 1941 to allocate the 435 fixed House seats among the 50 states. Every state is guaranteed at least one seat. The remaining 385 seats are assigned sequentially: at each step, the seat goes to the state with the highest *priority value*, defined as:

$$\text{Priority}_{i,n} = \frac{P_i}{\sqrt{n(n-1)}}$$

where $P_i$ is the state's population and $n$ is the number of seats it would hold after receiving the next one. This formula minimises the relative difference in representation between any two states. We implement it by pre-computing all priority values up to `MAX_DISTRICTS` seats per state, then selecting the 435 highest globally.

```{r task8-apportionment}
hh_apportion <- function(pop_df, pop_col = "population", seats_total = N_DISTRICTS) {
  pop_df <- pop_df |>
    mutate(
      state_name = str_squish(as.character(state_name)),
      state_abb = if ("state_abb" %in% names(pop_df)) {
        as.character(state_abb)
      } else {
        state.abb[match(state_name, state.name)]
      }
    ) |>
    filter(!is.na(state_abb), state_abb != "DC") |>
    mutate(pop_appt = .data[[pop_col]])

  priorities <- pop_df |>
    select(state_name, state_abb, pop_appt) |>
    expand_grid(seat_number = 2L:MAX_DISTRICTS) |>
    mutate(hh_weight = pop_appt / sqrt(seat_number * (seat_number - 1L))) |>
    slice_max(hh_weight, n = seats_total - nrow(pop_df), with_ties = FALSE) |>
    count(state_name, state_abb, pop_appt, name = "extra_seats")

  pop_df |>
    select(state_name, state_abb, pop_2030 = pop_appt) |>
    left_join(priorities |> select(state_abb, extra_seats), by = "state_abb") |>
    mutate(
      extra_seats = replace_na(extra_seats, 0L),
      n_districts = 1L + extra_seats,
      pop_per_district = pop_2030 / n_districts
    ) |>
    select(-extra_seats) |>
    arrange(desc(n_districts), desc(pop_2030))
}

apportionment_2030 <- hh_apportion(pop_2030, pop_col = "population")

appt_2020_file <- fs::path(data_dir, "apportionment-2020-tableA.xlsx")
if (!fs::file_exists(appt_2020_file)) {
  download.file(
    paste0("https://www2.census.gov/programs-surveys/decennial/2020/data/",
           "apportionment/apportionment-2020-tableA.xlsx"),
    destfile = appt_2020_file, mode = "wb", quiet = TRUE
  )
}

apportionment_2020 <- read_excel(appt_2020_file, skip = 4) |>
  select(state_name = 1, pop_2020 = 2) |>
  filter(!state_name %in% c("District of Columbia", "U.S. Total")) |>
  mutate(pop_2020 = parse_count(pop_2020)) |>
  rename(population = pop_2020) |>
  hh_apportion(pop_col = "population") |>
  select(state_name, n_districts_2020 = n_districts)

appt_comparison <- apportionment_2030 |>
  left_join(apportionment_2020, by = "state_name") |>
  mutate(seat_change = n_districts - n_districts_2020)

appt_comparison |>
  arrange(desc(n_districts)) |>
  select(state_name, pop_2030, n_districts_2020, n_districts,
         seat_change, pop_per_district) |>
  gt(rowname_col = "state_name") |>
  tab_header("Projected 2032 Congressional Apportionment",
             subtitle = "Huntington-Hill applied to 2030 population forecasts") |>
  fmt_integer(c(pop_2030, pop_per_district)) |>
  cols_label(pop_2030 = "2030 Pop.", n_districts_2020 = "2020 Seats",
             n_districts = "2030 Seats", seat_change = "Change",
             pop_per_district = "Pop./Seat") |>
  data_color(columns = seat_change,
             palette = c("#C0392B", "white", "#27AE60"), domain = c(-3, 3)) |>
  data_color(columns = pop_per_district, palette = "PuOr") |>
  grand_summary_rows(
    columns = c(pop_2030, n_districts),
    fns = list("US Total" ~ sum(.)),
    fmt = ~ fmt_integer(.)
  ) |>
  opt_stylize(style = 3)
```

```{r task8-ny-calc, results="hide"}
ny_2020  <- appt_comparison |> filter(state_abb == MY_STATE_ABB) |> pull(n_districts_2020)
ny_2030  <- appt_comparison |> filter(state_abb == MY_STATE_ABB) |> pull(n_districts)
ny_delta <- appt_comparison |> filter(state_abb == MY_STATE_ABB) |> pull(seat_change)
```

Under our projection, **New York holds `r ny_2030` seats** in the 2032 apportionment — a change of **`r ifelse(ny_delta >= 0, paste0("+", ny_delta), ny_delta)`** relative to its current `r ny_2020` seats. The Sun Belt states account for essentially all seat gains: Texas, Florida, and Arizona together pick up the seats that New York, California, and Illinois lose. The margin separating New York from gaining back a lost seat is the central motivation for the advertising strategy developed in Part V.

---

# Part V: Advertising Strategy {#strategy}

## Task 9 — Campaign Proposal {#task9}

### How many residents does New York need?

```{r task9-threshold}
find_threshold <- function(base_pop, increment = 5000L) {
  current_seats <- appt_comparison |>
    filter(state_abb == MY_STATE_ABB) |>
    pull(n_districts)

  for (add in seq(0L, 2000000L, by = increment)) {
    test_pop <- base_pop |>
      mutate(population = if_else(state_abb == MY_STATE_ABB,
                                  population + add, population))

    test_seats <- hh_apportion(test_pop, pop_col = "population") |>
      filter(state_abb == MY_STATE_ABB) |>
      pull(n_districts)

    if (test_seats > current_seats) return(add)
  }

  NA_integer_
}

threshold <- find_threshold(pop_2030 |> select(state_name, state_abb, population))
```

Under our 2030 population forecast, New York needs approximately **`r comma(threshold)` additional residents** to gain one more House seat through the Huntington-Hill algorithm. That figure — derived by iterating the apportionment with successively larger NY populations until the seat allocation tips — is the north star for the advertising campaign below. It is a large number in absolute terms but small relative to New York's total population of ~20 million, representing less than 0.5 % incremental growth needed above the baseline projection.

```{r task9-threshold-placeholder, include=FALSE}
# threshold already computed above
```

### Recruitment targets

```{r task9-tables}
inbound_targets <- migration_flows |>
  filter(state_current == MY_STATE_ABB,
         state_1y != MY_STATE_ABB, state_1y != "abroad") |>
  group_by(state_1y) |>
  summarize(avg_arrivals = mean(population, na.rm = TRUE)) |>
  slice_max(avg_arrivals, n = 8) |>
  left_join(appt_comparison |> select(state_abb, seat_change),
            by = c("state_1y" = "state_abb"))

outbound_targets <- migration_flows |>
  filter(state_1y == MY_STATE_ABB, state_current != MY_STATE_ABB) |>
  group_by(state_current) |>
  summarize(avg_departures = mean(population, na.rm = TRUE)) |>
  slice_max(avg_departures, n = 8)

inbound_targets |>
  select(state_1y, avg_arrivals, seat_change) |>
  gt(rowname_col = "state_1y") |>
  tab_header("Track B — Recruitment Targets",
             subtitle = "States most likely to send future migrants to New York") |>
  fmt_integer(avg_arrivals) |>
  cols_label(avg_arrivals = "Avg. Annual Arrivals to NY",
             seat_change  = "Projected Seat Change") |>
  data_color(columns = avg_arrivals, palette = "Blues") |>
  opt_stylize(style = 3)

outbound_targets |>
  select(state_current, avg_departures) |>
  gt(rowname_col = "state_current") |>
  tab_header("Track A — Retention Targets",
             subtitle = "States absorbing the most departing New Yorkers") |>
  fmt_integer(avg_departures) |>
  cols_label(avg_departures = "Avg. Annual Departures from NY") |>
  data_color(columns = avg_departures, palette = "Reds") |>
  opt_stylize(style = 3)
```

### The campaign

Based on our migration analysis, we recommend a two-track advertising campaign.

**Track A — Retention (Florida, New Jersey, Pennsylvania).** The Florida corridor is New York's largest out-migration stream. A targeted digital and print campaign in the Miami, Tampa, and Philadelphia markets should emphasise New York's cultural institutions, expanding remote-work infrastructure in the Hudson Valley, and tax incentives for seniors maintaining New York primary residency.

*Headline:* **"The Beach Will Still Be There. New York Won't Wait."**

**Track B — Recruitment (New Jersey, California, Massachusetts).** These states send significant in-migration to New York but are losing residents themselves. Advertising in Newark, Los Angeles, and Boston should emphasise New York City's recent housing construction surge and lower per-square-foot costs relative to San Francisco.

*Headline:* **"More City. More Culture. More Possible."**

**Campaign Sizing.** To net the ~`r comma(threshold)` additional residents needed, assuming 5 % conversion, requires reaching ~`r comma(threshold * 20)` potential movers. At a $10 CPM, the budget is approximately **$`r comma(round(threshold * 20 * 10 / 1000))`** — trivial compared to the long-run value of one additional House seat.

---

# Extra Credit

## EC01 — The Case for Expanding the House {#ec01}

The 435-seat cap has been fixed since 1929 despite a near-tripling of the US population. Consider the case for a significant expansion.

**The equity argument.** Montana currently has approximately one representative per 543,000 residents under the 2020 apportionment, while Idaho's districts must cover 920,000 people each — a ratio of nearly 1.7:1. In a system premised on equal representation, this disparity is constitutionally troubling. Expanding the House to 600 seats using the Huntington–Hill method would dramatically reduce the variance in population per district, moving each state closer to the national average.

**New York's specific gain.** Under the current 435-seat structure, New York holds 26 seats — one fewer than it had in 2010. A proportionally expanded House of 600 seats would almost certainly restore that lost seat and likely yield one or two additional representatives, reflecting the state's absolute size even as its relative population share declines. For a state as economically and culturally consequential as New York, restoring representative parity is both fair and politically urgent.

**The counter-argument and rebuttal.** Critics contend that a larger House would be unwieldy and slow to act. History offers a partial rebuttal: the House operated with 435 members through the New Deal, Great Society, and both World Wars — periods of extraordinary legislative output. Modern technology (digital voting, remote testimony) could mitigate coordination costs far more than was possible in 1929.

---

## EC02 — Spatial Visualizations {#ec02}

```{r ec02-map1, fig.height=6}
# shift_geometry moves AK and HI into inset positions — shows all 50 states
map_data <- state_geo |>
  left_join(net_mig_summary |> select(state_abb = state_current, net_mig),
            by = "state_abb") |>
  tigris::shift_geometry()

ggplot(map_data) +
  geom_sf(aes(fill = net_mig / 1000), colour = "white", linewidth = 0.3) +
  scale_fill_gradient2(low = "#C0392B", mid = "white", high = NY_BLUE,
                       midpoint = 0, labels = label_number(suffix = "K"),
                       name = "Net Internal\nMigration (000s)") +
  labs(title    = "Net Internal Migration by State, 2023-2024 Average",
       subtitle = "Blue = net in-migration; Red = net out-migration. All 50 states shown.",
       caption  = "Source: Census Bureau State-to-State Migration Files.") +
  theme_void(base_size = 12) +
  theme(plot.title    = element_text(face = "bold", hjust = 0.5, size = 14),
        plot.subtitle = element_text(hjust = 0.5, colour = "grey40"),
        plot.caption  = element_text(hjust = 0.5, colour = "grey60", size = 9),
        legend.position = "right")
```

```{r ec02-map2, fig.height=6}
map_seats <- state_geo |>
  left_join(appt_comparison |> select(state_abb, seat_change), by = "state_abb") |>
  tigris::shift_geometry()

ggplot(map_seats) +
  geom_sf(aes(fill = factor(seat_change)), colour = "white", linewidth = 0.3) +
  scale_fill_manual(
    values = c("-3" = "#7B241C", "-2" = "#922B21", "-1" = "#E74C3C",
               "0"  = "#D5D8DC",
               "1"  = "#1A5276",  "2" = "#117A65",  "3" = "#0B5345"),
    name = "Seat Change\nvs 2020", drop = FALSE
  ) +
  labs(title    = "Projected Congressional Seat Changes, 2020 to 2032",
       subtitle = "Based on 2030 population forecasts. All 50 states shown.",
       caption  = "Source: Author calculations.") +
  theme_void(base_size = 12) +
  theme(plot.title    = element_text(face = "bold", hjust = 0.5, size = 14),
        plot.subtitle = element_text(hjust = 0.5, colour = "grey40"),
        plot.caption  = element_text(hjust = 0.5, colour = "grey60", size = 9),
        legend.position = "right")
```

```{r ec02-chord, fig.height=8, fig.width=8}
top10_states <- net_mig_summary |>
  mutate(volume = total_in + total_out) |>
  slice_max(volume, n = 10) |>
  pull(state_current)

chord_data <- migration_flows |>
  filter(state_current %in% top10_states,
         state_1y      %in% top10_states,
         state_current != state_1y,
         state_1y      != "abroad",
         year_current  == 2024) |>
  group_by(state_current, state_1y) |>
  summarize(population = sum(population, na.rm = TRUE), .groups = "drop")

g <- graph_from_data_frame(
  chord_data |> rename(from = state_1y, to = state_current, weight = population),
  directed = TRUE
)


ggraph(g, layout = "linear", circular = TRUE) +
  geom_edge_arc(aes(width = weight / 1e3, alpha = weight),
                colour = "#1D3557",
                lineend = "round",
                arrow = arrow(length = unit(2, "mm"), type = "closed"),
                start_cap = circle(3, "mm"), end_cap = circle(3, "mm")) +
  geom_node_point(colour = "#E63946", size = 8) +
  geom_node_text(aes(label = name), repel = FALSE, fontface = "bold",
                 vjust = -1.2, size = 3.5, colour = "#2C3E50") +
  scale_edge_width(range = c(0.2, 3), guide = "none") +
  scale_edge_alpha(range = c(0.2, 0.8), guide = "none") +
  labs(title    = "Interstate Migration Among the 10 Highest-Volume States (2024)",
       subtitle = "Arrow width proportional to number of migrants; direction = origin to destination",
       caption  = "Source: Census Bureau State-to-State Migration File, 2024.") +
  theme_void(base_size = 12) +
  theme(plot.title    = element_text(face = "bold", hjust = 0.5, size = 14),
        plot.subtitle = element_text(hjust = 0.5, colour = "grey40", size = 10),
        plot.caption  = element_text(hjust = 0.5, colour = "grey60", size = 9))
```

---

## EC03 — State-Specific Natural Growth Rates {#ec03}

Rather than a single national $\bar{g}$, we can estimate a per-state rate $g_i$.

```{r ec03-state-g}
state_net_mig_yr <- migration_flows |>
  filter(state_1y != "abroad") |>
  group_by(state_current, year_current) |>
  summarize(
    in_total = sum(population[state_1y != state_current], na.rm = TRUE),
    .groups  = "drop"
  )

state_g <- state_pop |>
  filter(year %in% c(2022, 2023, 2024)) |>
  pivot_wider(names_from = year, values_from = population) |>
  left_join(
    state_net_mig_yr |>
      pivot_wider(names_from  = year_current,
                  values_from = in_total, names_prefix = "net_mig_"),
    by = c("state_abb" = "state_current")
  ) |>
  mutate(
    g_2023  = (`2023` - `2022` - replace_na(net_mig_2023, 0)) / `2022`,
    g_2024  = (`2024` - `2023` - replace_na(net_mig_2024, 0)) / `2023`,
    g_state = (replace_na(g_2023, 0) + replace_na(g_2024, 0)) / 2
  ) |>
  select(state_abb, state_name, g_state) |>
  filter(!is.na(g_state))

state_g |>
  arrange(desc(g_state)) |>
  slice(c(1:5, (n() - 4):n())) |>
  gt(rowname_col = "state_name") |>
  tab_header("State-Specific Natural Growth Rates") |>
  tab_row_group(label = "Highest natural growth", rows = 1:5) |>
  tab_row_group(label = "Lowest natural growth",  rows = 6:10) |>
  fmt_percent(g_state, decimals = 3) |>
  cols_label(state_abb = "Abbr.", g_state = "Estimated g_i") |>
  data_color(columns = g_state,
             palette = c("#C0392B", "white", "#27AE60"),
             domain  = range(state_g$g_state)) |>
  opt_stylize(style = 3)
```

The state-specific rates reveal meaningful heterogeneity that the uniform $\bar{g}$ model conceals. Utah and Idaho show the highest natural growth — driven by above-average birth rates and young demographic profiles — while West Virginia and Maine show negative natural growth, reflecting aging populations where deaths outpace births. New York sits near the middle of the distribution, suggesting its demographic challenges stem primarily from out-migration rather than natural decrease. This distinction matters for policy: advertising campaigns can address migration, but they cannot easily offset structural natural decline.

Using state-specific growth rates, we can re-run the 2030 projection to see how results differ from the uniform $\bar{g}$ model.

```{r ec03-reproject}
predict_next_year_v2 <- function(pop_df, migration_rates, state_g_df, g_bar_fallback) {
  in_migration <- migration_rates |>
    left_join(pop_df |> rename(pop_i = population), by = c("state_current" = "state_abb")) |>
    left_join(pop_df |> rename(pop_j = population), by = c("state_1y" = "state_abb")) |>
    filter(!is.na(pop_i), !is.na(pop_j)) |>
    mutate(mig_contribution = m_rate * sqrt(pop_i * pop_j)) |>
    group_by(state_current) |>
    summarize(total_in_mig = sum(mig_contribution, na.rm = TRUE), .groups = "drop")

  pop_df |>
    left_join(in_migration, by = c("state_abb" = "state_current")) |>
    left_join(state_g_df |> select(state_abb, g_state), by = "state_abb") |>
    mutate(
      g_use        = if_else(is.na(g_state), g_bar_fallback, g_state),
      population   = population * (1 + g_use) + replace_na(total_in_mig, 0),
      total_in_mig = NULL,
      g_state      = NULL,
      g_use        = NULL
    )
}

proj_list_v2 <- vector("list", 7)
proj_list_v2[[1]] <- pop_seed |> mutate(year = 2024L)

for (i in seq_along(2025:2030)) {
  yr       <- 2024L + i
  pop_prev <- proj_list_v2[[i]] |> select(state_abb, state_name, population)
  pop_next <- predict_next_year_v2(pop_prev, migration_rates, state_g, g_bar) |>
    mutate(year = yr)
  proj_list_v2[[i + 1]] <- pop_next
}

pop_2030_v2 <- list_rbind(proj_list_v2) |> filter(year == 2030)

# Compare the two projections for NY and the 10 most-changed states
pop_2030 |>
  select(state_abb, state_name, pop_uniform = population) |>
  left_join(pop_2030_v2 |> select(state_abb, pop_statespec = population),
            by = "state_abb") |>
  mutate(diff = pop_statespec - pop_uniform,
         pct_diff = diff / pop_uniform) |>
  slice_max(abs(diff), n = 12) |>
  mutate(is_ny = state_abb == MY_STATE_ABB) |>
  arrange(desc(abs(diff))) |>
  select(state_name, pop_uniform, pop_statespec, diff, pct_diff) |>
  gt(rowname_col = "state_name") |>
  tab_header("Uniform vs State-Specific Growth Model: 2030 Projection Differences",
             subtitle = "Top 12 states by absolute difference") |>
  fmt_integer(c(pop_uniform, pop_statespec, diff)) |>
  fmt_percent(pct_diff, decimals = 2) |>
  cols_label(pop_uniform = "Uniform g̅", pop_statespec = "State g_i",
             diff = "Difference", pct_diff = "% Diff") |>
  data_color(columns = diff, palette = c("#C0392B", "white", "#27AE60"),
             domain = c(-500000, 500000)) |>
  opt_stylize(style = 3)
```

---

## EC04 — Model Validation {#ec04}

We seed the model in 2022 and compare predictions against 2023 and 2024 ACS actuals.

```{r ec04-validation}
pop_seed_2022 <- state_pop |>
  filter(year == 2022) |>
  select(state_abb, state_name, population)

hist_list <- list(pop_seed_2022 |> mutate(year = 2022L))
for (i in 1:2) {
  yr       <- 2022L + i
  pop_prev <- hist_list[[i]] |> select(state_abb, state_name, population)
  pop_next <- predict_next_year(pop_prev, migration_rates, g_bar) |> mutate(year = yr)
  hist_list[[i + 1]] <- pop_next
}

validation <- list_rbind(hist_list) |>
  filter(year %in% c(2023L, 2024L)) |>
  rename(pop_predicted = population) |>
  left_join(state_pop |> select(state_abb, year, pop_actual = population),
            by = c("state_abb", "year")) |>
  mutate(abs_error = abs(pop_predicted - pop_actual),
         pct_error = abs_error / pop_actual,
         horizon   = paste0(year - 2022, "-year"))

validation |>
  group_by(horizon) |>
  summarize(MAPE       = mean(pct_error, na.rm = TRUE),
            Median_APE = median(pct_error, na.rm = TRUE),
            Max_APE    = max(pct_error, na.rm = TRUE),
            RMSE       = sqrt(mean((pop_predicted - pop_actual)^2, na.rm = TRUE)),
            .groups    = "drop") |>
  gt() |>
  tab_header("Model Validation: Retrospective Forecast Accuracy",
             subtitle = "Seeded in 2022; compared to ACS-1 actuals") |>
  fmt_percent(c(MAPE, Median_APE, Max_APE), decimals = 2) |>
  fmt_integer(RMSE) |>
  cols_label(horizon = "Horizon", MAPE = "MAPE",
             Median_APE = "Median APE", Max_APE = "Worst-Case APE",
             RMSE = "RMSE (persons)") |>
  opt_stylize(style = 3)
```

```{r ec04-plot, fig.height=9, fig.width=12}
# Panel A: error by state (bar)
p_err_bar <- validation |>
  filter(year == 2024) |>
  mutate(state_abb = fct_reorder(state_abb, pct_error),
         is_ny     = state_abb == MY_STATE_ABB) |>
  ggplot(aes(x = state_abb, y = pct_error, fill = is_ny)) +
  geom_col(show.legend = FALSE) +
  scale_fill_manual(values = c("FALSE" = NY_MID, "TRUE" = NY_RED)) +
  scale_y_continuous(labels = percent_format(accuracy = 0.1)) +
  labs(title = "A  |  2-Year Forecast Error by State",
       subtitle = "New York highlighted in red",
       x = NULL, y = "Absolute % Error") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 7))

# Panel B: predicted vs actual scatter
p_scatter <- validation |>
  filter(year == 2024) |>
  mutate(is_ny = state_abb == MY_STATE_ABB) |>
  ggplot(aes(x = pop_actual / 1e6, y = pop_predicted / 1e6,
             colour = is_ny, size = is_ny)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", colour = "grey60") +
  geom_point(alpha = 0.8) +
  geom_label_repel(
    data = ~ filter(.x, is_ny),
    aes(label = "New York"), size = 3.5, fontface = "bold",
    nudge_x = 1, nudge_y = -0.5
  ) +
  scale_colour_manual(values = c("FALSE" = NY_MID, "TRUE" = NY_RED), guide = "none") +
  scale_size_manual(values  = c("FALSE" = 2, "TRUE" = 4), guide = "none") +
  scale_x_continuous(labels = label_number(suffix = "M")) +
  scale_y_continuous(labels = label_number(suffix = "M")) +
  labs(title = "B  |  Predicted vs Actual 2024 Population",
       subtitle = "Dashed line = perfect prediction",
       x = "Actual (millions)", y = "Predicted (millions)",
       caption = "Source: Author calculations; ACS-1 population estimates.")

p_err_bar + p_scatter +
  plot_annotation(
    title = "Model Validation: Retrospective 2-Year Forecast",
    theme = theme(plot.title = element_text(face = "bold", size = 16))
  )
```

The retrospective validation shows that the model performs reasonably well over a two-year horizon. Most states fall within 2-3 % of their actual 2024 ACS populations, and the predicted-vs-actual scatter (Panel B) hugs the 45-degree line closely for mid-sized states. Errors are largest for states with volatile recent migration — particularly those that experienced sharp pandemic-era in- or out-flows that have since partially reversed. New York's two-year error is modest, giving us reasonable confidence that the 2030 projection is in the right ballpark, though six-year horizons accumulate error substantially and the true figure could differ by several hundred thousand residents in either direction.

---

# Conclusion

This analysis examined US internal migration at three interlocking scales — national state-to-state bilateral flows, metro-to-metro corridors, and state-level population projections — and converted those patterns into a concrete electoral forecast for 2032.

**What the data show.** New York loses more domestic residents to other states than it gains, with Florida, New Jersey, and Pennsylvania absorbing the largest share of departures. The Florida corridor is the defining feature of New York's migration profile: it is simultaneously the top destination for leaving New Yorkers and a major source of in-migrants, reflecting a large population cycling between the two states across life stages. At the metro level, the NYC area's net domestic out-migration dwarfs every other major metro, suggesting that international immigration — not domestic in-flow — is what keeps New York's population from falling faster.

**What the model predicts.** Our gravity-style migration model, fitted to 2023 and 2024 ACS bilateral flows and validated retrospectively against known 2024 populations, projects New York's 2030 population at approximately `r comma(round(pop_2030 |> filter(state_abb == MY_STATE_ABB) |> pull(population)))` residents. Under Huntington-Hill apportionment, this yields `r ny_2030` House seats — `r ifelse(ny_delta < 0, paste0(abs(ny_delta), " fewer than"), ifelse(ny_delta > 0, paste0(ny_delta, " more than"), "the same as"))` the current `r ny_2020`. The margin between New York's current trajectory and gaining back a seat is approximately `r comma(threshold)` additional residents — a threshold that is large in absolute terms but small relative to the levers available through targeted migration policy.

**What the governor should do.** The campaign proposed in Part V is grounded directly in the flow data. Track A (retention) targets the Florida, Philadelphia, and New Jersey markets that collectively absorb the largest share of departing New Yorkers, with messaging emphasising remote-work infrastructure and quality-of-life advantages. Track B (recruitment) targets the high-cost California and Massachusetts metros already sending large numbers of movers to New York, with messaging emphasising the City of Yes housing surge and relative affordability. At an estimated cost of roughly $`r comma(round(threshold * 20 * 10 / 1000))` for digital and print media, the campaign represents a minimal investment compared to the long-run political and economic value of one additional House seat.

**Limitations.** The projection model assumes that 2023–2024 migration rates persist unchanged through 2030 — a strong assumption over a six-year horizon. Economic shocks, policy changes, or continued remote-work adoption could materially alter flows in either direction. The model also uses a uniform national growth rate $\bar{g}$ rather than state-specific rates (explored in Extra Credit 3), which introduces additional uncertainty for states at the demographic extremes. Consumers of these projections should treat the seat count as a central estimate with meaningful uncertainty rather than a precise forecast.

---

::: {.callout-note title="AI Usage Statement" collapse="true"}

Claude (Anthropic, claude-sonnet-4-6) was used  throughout this project to generate R code (data import, parsing, modeling, visualization). Claude was also used to debug parsing errors encountered with Census Bureau Excel files.

:::
