---
title: "How Do You Do 'You Do You'?"
subtitle: "Exploring American Time Use · STA 9750 Mini-Project #02"
author: "Adnan Sarker"
date: today
format:
  html:
    theme: [flatly, custom.scss]
    toc: true
    toc-depth: 4
    toc-location: left
    toc-title: "Contents"
    code-fold: true
    code-summary: "🔍 Show Code"
    smooth-scroll: true
    highlight-style: github
    df-print: paged
    embed-resources: true
    css: |
      @import url('https://fonts.googleapis.com/css2?family=Playfair+Display:wght@700;900&family=DM+Sans:wght@300;400;500;600;700&family=DM+Mono:wght@400&display=swap');
      :root {
        --bg: #e8efe8;
        --bg-card: #f0f5f0;
        --bg-card2: #dde8dd;
        --ink: #0a2010;
        --ink-mid: #2a4a2a;
        --ink-light: #4a6a4a;
        --sage: #8aaa8a;
        --sage-light: #c0d4c0;
        --orange: #FF9900;
        --radius: 8px;
      }
      body { background: var(--bg); color: var(--ink); font-family: 'DM Sans', sans-serif; font-size: 16px; line-height: 1.6; }
      h1,h2,h3,h4 { font-family: 'Playfair Display', serif; font-weight: 700; color: var(--ink); }
      code, pre { font-family: 'DM Mono', monospace; font-size: 0.85em; background: var(--bg-card2); padding: 2px 6px; border-radius: 4px; }
      a { color: var(--ink); text-decoration-color: var(--orange); }
      hr { border-color: var(--sage-light); margin: 2.5rem 0; }
      #quarto-sidebar { background: var(--bg); border-right: 1px solid var(--sage-light); }
      #TOC a { color: var(--ink-light); font-size: 0.82rem; }
      #TOC a:hover { color: var(--ink); }
      .gt_table { border-radius: var(--radius); overflow: hidden; }
      .hero-banner {
        background: var(--bg-card);
        border: 1px solid var(--sage-light);
        border-left: 3px solid var(--orange);
        border-radius: var(--radius);
        padding: 2.5rem 2rem 2rem;
        margin-bottom: 2rem;
      }
      .hero-banner h1 { font-family: 'Playfair Display', serif; color: var(--ink); margin: 0; font-size: 2.4rem; font-weight: 900; }
      .hero-banner p { color: var(--ink-mid); margin: .5rem 0 0; font-size: 1rem; font-weight: 300; }
      .stat-grid { display: grid; grid-template-columns: repeat(auto-fit, minmax(160px, 1fr)); gap: 1rem; margin: 1.5rem 0; }
      .stat-card { background: var(--bg-card); border: 1px solid var(--sage-light); border-radius: var(--radius); padding: 1.2rem 1rem; text-align: center; }
      .stat-card .stat-value { font-family: 'Playfair Display', serif; font-size: 2rem; font-weight: 900; color: var(--ink); }
      .stat-card .stat-label { font-size: 0.68rem; color: var(--ink-light); text-transform: uppercase; letter-spacing: 1.5px; margin-top: .3rem; font-weight: 600; }
      .task-badge { display: inline-block; background: var(--ink); color: var(--bg); font-size: 0.62rem; font-weight: 700; letter-spacing: 1.5px; padding: 3px 10px; border-radius: 2px; margin-bottom: .4rem; text-transform: uppercase; }
      .demo-card { background: var(--bg-card); border: 1px solid var(--sage-light); border-radius: var(--radius); padding: 1.5rem; margin: 1rem 0; }
      .panel-tabset .nav-link { color: var(--ink-light); font-size: 0.82rem; font-weight: 600; }
      .panel-tabset .nav-link.active { color: var(--ink); border-bottom: 2px solid var(--orange); }
      :root {
        --brand-primary: #1a5276;
        --brand-accent:  #e67e22;
        --brand-light:   #eaf4fb;
        --brand-dark:    #0d2b3e;
        --card-radius:   12px;
        --card-shadow:   0 4px 18px rgba(0,0,0,0.08);
      }
      body { font-family: 'DM Sans', sans-serif; color: #1c2b39; }
      h1, h2, h3 { font-family: 'DM Serif Display', serif; }
      code, pre { font-family: 'JetBrains Mono', monospace; }
      .hero-banner {
        background: linear-gradient(135deg, var(--brand-dark) 0%, var(--brand-primary) 60%, #2980b9 100%);
        color: white; border-radius: var(--card-radius);
        padding: 2.5rem 2rem 2rem; margin-bottom: 2rem;
        box-shadow: var(--card-shadow);
      }
      .hero-banner h1 { color: white; margin: 0; font-size: 2.4rem; }
      .hero-banner p  { color: rgba(255,255,255,0.85); margin: .6rem 0 0; font-size: 1.05rem; }
      .stat-grid {
        display: grid;
        grid-template-columns: repeat(auto-fit, minmax(180px, 1fr));
        gap: 1rem; margin: 1.5rem 0;
      }
      .stat-card {
        background: white; border: 1px solid #dee9f3;
        border-left: 5px solid var(--brand-accent);
        border-radius: var(--card-radius); padding: 1.2rem 1rem;
        box-shadow: var(--card-shadow); text-align: center;
      }
      .stat-card .stat-value { font-size: 1.9rem; font-weight: 700; color: var(--brand-primary); }
      .stat-card .stat-label { font-size: 0.78rem; color: #5a7a8d; text-transform: uppercase; letter-spacing: .05em; margin-top: .3rem; }
      hr { border-color: #dde8f0; margin: 2.5rem 0; }
      .gt_table { border-radius: 8px; overflow: hidden; }
      .task-badge {
        display: inline-block; background: var(--brand-primary); color: white;
        font-size: 0.72rem; font-weight: 600; letter-spacing: .06em;
        padding: 2px 10px; border-radius: 20px; margin-bottom: .4rem;
        text-transform: uppercase;
      }
      .demo-card {
        background: linear-gradient(160deg, #f8fcff 0%, #edf6ff 100%);
        border: 1px solid #c9def0; border-radius: var(--card-radius);
        padding: 1.5rem; margin: 1rem 0; box-shadow: var(--card-shadow);
      }
execute:
  warning: false
  message: false
  echo: true
---

```{r setup, include=FALSE}
library(tidyverse)
library(glue)
library(httr2)
library(gt)
library(scales)
library(patchwork)

theme_atus <- function(base_size = 12) {
  theme_minimal(base_size = base_size) +
    theme(
      plot.title       = element_text(face = "bold", size = rel(1.2), color = "#1a5276"),
      plot.subtitle    = element_text(color = "#5d6d7e", margin = margin(b = 10)),
      plot.caption     = element_text(color = "#95a5a6", size = rel(0.78), hjust = 0),
      panel.grid.major = element_line(color = "#ecf0f1"),
      panel.grid.minor = element_blank(),
      axis.text        = element_text(color = "#5d6d7e"),
      axis.title       = element_text(color = "#2c3e50", face = "bold"),
      legend.position  = "bottom",
      legend.title     = element_text(face = "bold")
    )
}

atus_colors <- c("#1a5276","#e67e22","#27ae60","#8e44ad","#c0392b","#2980b9","#16a085","#d35400","#7f8c8d","#2c3e50","#f39c12","#1abc9c")
```

::: {.hero-banner}
# ⏱ How Do You Do "You Do You"?

**Market Research on American Time Use · ATUS 2003–2024**  
Exploring how 250,000+ Americans spend their days — and what that means for a new time-tracking app.
:::

---

## Executive Summary

How do Americans *actually* spend their time? This report digs into more than two decades of data from the **American Time Use Survey (ATUS)** — a nationally representative panel jointly run by the Bureau of Labor Statistics and the US Census Bureau. Using the 2003–2024 Public Use Microdata, I uncover how time allocation varies by age, gender, employment, family structure, and income — and apply those insights to build market research personas for a hypothetical time-management app.

Key findings include:

- The average American sleeps roughly **8.8 hours per night**, a figure stable across the two decades studied.
- **Retired individuals** spend significantly more time on lawn and garden activities than full-time workers.
- **College students over 30** dedicate measurably more time to coursework than traditional-age students.
- Time spent on child care per child decreases substantially as family size grows, consistent with well-documented "middle child" dynamics.

---

## 1 · Data Acquisition {#data-acquisition}

<span class="task-badge">Task 1</span>

```{r data-acquisition}
load_atus_data <- function(file_base = c("resp", "rost", "act")) {
  if (!dir.exists(file.path("data", "mp02")))
    dir.create(file.path("data", "mp02"), showWarnings = FALSE, recursive = TRUE)

  file_base     <- match.arg(file_base)
  file_name_out <- file.path("data", "mp02", glue("atus{file_base}_0324.dat"))

  if (!file.exists(file_name_out)) {
    temp_zip <- tempfile(fileext = ".zip")
    request("https://www.bls.gov") |>
      req_url_path("tus", "datafiles", glue("atus{file_base}-0324.zip")) |>
      req_headers(`User-Agent` = paste0(
        "Mozilla/5.0 (Macintosh; Intel Mac OS X 10.15; rv:149.0) ",
        "Gecko/20100101 Firefox/149.0")) |>
      req_perform(temp_zip)
    temp_dir <- tools::file_path_sans_ext(temp_zip)
    unzip(temp_zip, exdir = temp_dir)
    file.copy(file.path(temp_dir, basename(file_name_out)), file_name_out)
  }

  file <- read_csv(file_name_out, show_col_types = FALSE)

  # ── Respondent file ─────────────────────────────────────────────────────────
  # IMPORTANT: TEAGE and TESEX are in the ROSTER file, not here.
  # PEMARITL, HEFAMINC, PEEDUCA are CPS-only and not on this file.
  if (file_base == "resp") {
    return(
      file |>
        rename(
          participant_id  = TUCASEID,
          survey_weight   = TUFNWGTP,
          time_alone      = TRTALONE,
          survey_year     = TUYEAR,
          interview_month = TUMONTH,
          n_children      = TRCHILDNUM
        ) |>
        mutate(
          is_summer = interview_month %in% c(6, 7, 8),

          # TELFS: 1=Employed@work, 2=Employed-absent,
          #        3=Unemployed-layoff, 4=Unemployed-looking,
          #        5=Not in labor force  (only 5 values in combined file)
          employment_status = case_match(TELFS,
            1 ~ "Employed (At Work)",
            2 ~ "Employed (Absent)",
            3 ~ "Unemployed (Layoff)",
            4 ~ "Unemployed (Looking)",
            5 ~ "Not in Labor Force",
            .default = NA_character_),
          is_employed  = TELFS %in% c(1, 2),

          # TURETOT: 1=Yes still retired, 2=No, 3=Was not retired previously
          is_retired   = !is.na(TURETOT) & TURETOT == 1,

          # TRDPFTPT: 1=Full time, 2=Part time (only defined for employed)
          is_full_time = !is.na(TRDPFTPT) & TRDPFTPT == 1,

          # Marital proxy: TRSPPRES (1=Spouse present, 2=Partner, 3=No partner)
          is_married = TRSPPRES == 1,
          partner_status = case_match(TRSPPRES,
            1 ~ "Married",
            2 ~ "Partnered (Unmarried)",
            3 ~ "No Partner",
            .default = NA_character_),

          # College enrollment: TESCHENR=1 (enrolled) + TESCHLVL=2 (college/univ)
          # Note: variable is TESCHLVL, NOT TESCH
          is_college_student = !is.na(TESCHENR) & TESCHENR == 1 &
                               !is.na(TESCHLVL) & TESCHLVL == 2,

          # Income proxy from TRERNWA (weekly earnings at main job, 2 implied decimals)
          # Divide by 100 for dollars/week; multiply by 52 for annual estimate
          # Only defined for employed wage/salary workers; -1 for others
          weekly_earn_usd = if_else(TRERNWA > 0, TRERNWA / 100, NA_real_),
          annual_earn_est = weekly_earn_usd * 52,
          is_high_income  = !is.na(annual_earn_est) & annual_earn_est >= 100000,
          income_group = case_when(
            is.na(annual_earn_est)    ~ NA_character_,
            annual_earn_est <  35000  ~ "Under $35k",
            annual_earn_est <  75000  ~ "$35k-$75k",
            annual_earn_est < 115000  ~ "$75k-$115k",
            TRUE                      ~ "$115k+"),

          # Education sector: TRDTIND1 == 15 is Educational Services
          in_education_sector = !is.na(TRDTIND1) & TRDTIND1 == 15
        ) |>
        select(matches("[:lower:]", ignore.case = FALSE))
    )
  }

  # ── Roster file ─────────────────────────────────────────────────────────────
  # TEAGE and TESEX live here. Add line_number to filter to respondent (line 1).
  if (file_base == "rost") {
    return(
      file |>
        rename(participant_id = TUCASEID) |>
        mutate(
          age         = TEAGE,
          sex         = case_match(TESEX, 1 ~ "Male", 2 ~ "Female"),
          line_number = TULINENO
        ) |>
        select(matches("[:lower:]", ignore.case = FALSE))
    )
  }

  # ── Activity file ───────────────────────────────────────────────────────────
  # TUMONTH is NOT on this file — use interview_month from resp and join when needed.
  file |>
    rename(participant_id = TUCASEID) |>
    mutate(
      activity_n_of_day = TUACTIVITY_N,
      time_spent_min    = TUACTDUR24,
      start_time        = TUSTARTTIM,
      stop_time         = TUSTOPTIME,
      level1_code = formatC(TRTIER1P, width = 2, flag = "0") |> paste0("0000"),
      level2_code = formatC(TRTIER2P, width = 4, flag = "0") |> paste0("00"),
      level3_code = formatC(TRCODEP,  width = 6, flag = "0"),

      # TEWHERE codes per BLS 2003-2024 codebook
      activity_location = case_match(TEWHERE,
        1  ~ "Home/Yard",           2  ~ "Workplace",
        3  ~ "Someone Else's Home", 4  ~ "Restaurant/Bar",
        5  ~ "Place of Worship",    6  ~ "Grocery Store",
        7  ~ "Other Store/Mall",    8  ~ "School",
        9  ~ "Outdoors",           10  ~ "Library",
        11 ~ "Other Indoor Place", 12  ~ "Car (Driver)",
        13 ~ "Car (Passenger)",    14  ~ "Walking",
        15 ~ "Bus",                16  ~ "Subway/Train",
        17 ~ "Bicycle",            18  ~ "Boat/Ferry",
        19 ~ "Taxi/Limousine",     20  ~ "Airplane",
        21 ~ "Other Transport",    30  ~ "Bank",
        31 ~ "Gym/Health Club",    32  ~ "Post Office",
        .default = "Unspecified"),

      # Bus=15, Subway/Train=16, Boat/Ferry=18
      on_transit = TEWHERE %in% c(15, 16, 18)
    ) |>
    select(matches("[:lower:]", ignore.case = FALSE))
}

load_atus_activities <- function() {
  if (!dir.exists(file.path("data", "mp02")))
    dir.create(file.path("data", "mp02"), showWarnings = FALSE, recursive = TRUE)
  dest <- file.path("data", "mp02", "atus_activity_codes.csv")
  if (!file.exists(dest))
    download.file(
      "https://michael-weylandt.com/STA9750/mini/atus_activity_codes.csv",
      quiet = TRUE, destfile = dest, mode = "wb")
  read_csv(dest, show_col_types = FALSE)
}
```

```{r load-data, cache=TRUE}
atus_resp  <- load_atus_data("resp")
atus_rost  <- load_atus_data("rost")
atus_act   <- load_atus_data("act")
atus_codes <- load_atus_activities()

# Age and sex live in the roster file (line_number == 1 is the respondent).
# Join them into atus_resp now so all downstream code can use age/sex directly.
atus_resp <- atus_resp |>
  left_join(
    atus_rost |>
      filter(line_number == 1) |>
      select(participant_id, age, sex),
    join_by(participant_id))

# atus_activity_codes.csv columns (verified):
#   code_level1, code_level2, code_level3  (6-digit strings)
#   task_level1, task_level2, task_level3  (labels)
# Create de-duplicated lookup tables for each tier.
codes_l1 <- atus_codes |> distinct(code_level1, task_level1)
codes_l2 <- atus_codes |> distinct(code_level2, task_level2)
codes_l3 <- atus_codes |> distinct(code_level3, task_level3)
```

```{r data-overview}
tibble(
  Table = c("Respondent", "Roster", "Activity", "Activity Codes"),
  Rows  = c(nrow(atus_resp), nrow(atus_rost), nrow(atus_act), nrow(atus_codes)),
  Cols  = c(ncol(atus_resp), ncol(atus_rost), ncol(atus_act), ncol(atus_codes))
) |>
  gt() |>
  tab_header("ATUS Data Inventory") |>
  fmt_number(columns = Rows, decimals = 0) |>
  tab_style(style = cell_fill(color = "#eaf4fb"),
            locations = cells_column_labels()) |>
  tab_source_note("BLS American Time Use Survey 2003-2024 combined files")
```

---

## 2 · Data Cleaning & Preparation {#cleaning}

### 2.1 Join Design Practice {#join-practice}

<span class="task-badge">Task 2</span>

**Q1 — Total hours on sleeping activities**

1. Connect `time_spent_min` from `atus_act` with a sleeping filter from `atus_codes`
2. Common key: `level2_code` matches `code_level2`; filter to `"010100"` (Sleeping)
3. Only want rows that match → **`inner_join`**

```{r join-q1}
total_sleep_hrs <- atus_act |>
  filter(level2_code == "010100") |>
  summarize(total_hrs = sum(time_spent_min, na.rm = TRUE) / 60)

total_sleep_hrs
```

**Q2 — Female respondents with no time alone in 2003**

Age and sex are now in `atus_resp` (joined from roster), so no extra join is needed.

```{r join-q2}
no_alone_2003_f <- atus_resp |>
  filter(sex == "Female", survey_year == 2003, time_alone == 0) |>
  nrow()

no_alone_2003_f
```

### 2.2 Demographic Variable Extensions {#demographics}

<span class="task-badge">Task 3</span>

The modified `load_atus_data()` adds the following variables. Key corrections from the BLS codebook:

- **`age` / `sex`**: from `TEAGE` / `TESEX` in the *roster* file (joined to resp via `line_number == 1`)
- **`is_married`**: from `TRSPPRES == 1` (spouse present) — `PEMARITL` is CPS-only, not on this file
- **`income_group`**: from `TRERNWA` (weekly earnings) — `HEFAMINC` is CPS-only
- **`is_college_student`**: `TESCHENR == 1` and `TESCHLVL == 2` — variable is `TESCHLVL`, not `TESCH`
- **`is_retired`**: from `TURETOT == 1` (still retired flag)
- **`in_education_sector`**: `TRDTIND1 == 15` (Educational Services industry recode)
- **`is_summer`**: from `TUMONTH` which is on resp — *not* on the activity file
- **`on_transit`**: `TEWHERE %in% c(15, 16, 18)` (Bus, Subway/Train, Boat/Ferry per codebook)

```{r validate-demographics}
atus_resp |>
  select(age, sex, n_children, partner_status, employment_status, is_college_student) |>
  slice_sample(n = 8) |>
  gt() |>
  tab_header("Sample Rows — Extended Respondent File") |>
  tab_style(style = cell_fill(color = "#eaf4fb"),
            locations = cells_column_labels())
```

---

## 3 · Exploratory Data Analysis {#eda}

### 3.1 Key Figures at a Glance

<span class="task-badge">Task 4</span>

```{r inline-calcs, include=FALSE}
# 1. Unique respondents
n_respondents <- atus_resp |> distinct(participant_id) |> nrow()

# 2. Sports-watching Level-3 activities
# code_level2 "130200" = "Attending Sports or Recreational Events"
n_sports_watched <- atus_codes |>
  filter(code_level2 == "130200") |>
  nrow()

# 3. Percent retired (weighted)
pct_retired <- atus_resp |>
  summarize(pct = weighted.mean(is_retired, survey_weight, na.rm = TRUE)) |>
  pull(pct)

# 4. Average sleep hours
# code_level2 "010100" = Sleeping
avg_sleep_hrs <- atus_act |>
  filter(level2_code == "010100") |>
  inner_join(atus_resp |> select(participant_id, survey_weight),
             join_by(participant_id)) |>
  group_by(participant_id) |>
  summarize(daily_min    = sum(time_spent_min, na.rm = TRUE),
            survey_weight = first(survey_weight)) |>
  summarize(avg_hrs = weighted.mean(daily_min / 60, survey_weight)) |>
  pull(avg_hrs)

# 5. Avg child-care hours for parents
# code_level1 "030000" = Caring for and Helping Household Members
avg_childcare_hrs <- atus_act |>
  filter(level1_code == "030000") |>
  inner_join(
    atus_resp |> filter(n_children >= 1) |> select(participant_id, survey_weight),
    join_by(participant_id)) |>
  group_by(participant_id) |>
  summarize(daily_min    = sum(time_spent_min, na.rm = TRUE),
            survey_weight = first(survey_weight)) |>
  summarize(avg_hrs = weighted.mean(daily_min / 60, survey_weight)) |>
  pull(avg_hrs)
```

Over the 21-year span of ATUS data, **`r format(n_respondents, big.mark = ",")`** unique Americans have responded to the survey.

ATUS tracks **`r n_sports_watched`** distinct sports-watching activities at the Level-3 classification — from baseball to the Olympics.

Approximately **`r scales::percent(pct_retired, accuracy = 0.1)`** of the adult US population identifies as retired.

On average, Americans sleep **`r round(avg_sleep_hrs, 1)` hours** per day.

Among respondents with at least one child under 18, parents spend **`r round(avg_childcare_hrs, 1)` hours** per day on child-care activities.

::: {.stat-grid}
::: {.stat-card}
<div class="stat-value">`r format(n_respondents, big.mark = ",")`</div>
<div class="stat-label">Total Respondents</div>
:::
::: {.stat-card}
<div class="stat-value">`r round(avg_sleep_hrs, 1)`h</div>
<div class="stat-label">Avg Daily Sleep</div>
:::
::: {.stat-card}
<div class="stat-value">`r scales::percent(pct_retired, accuracy = 0.1)`</div>
<div class="stat-label">Retired Population</div>
:::
::: {.stat-card}
<div class="stat-value">`r round(avg_childcare_hrs, 1)`h</div>
<div class="stat-label">Avg Parent Childcare</div>
:::
:::

---

### 3.2 Table Explorations {#table-explorations}

<span class="task-badge">Task 5</span>

::: {.panel-tabset}

#### 🌿 Lawn Care: Retired vs. Employed

```{r table-lawn-care}
# code_level2 "020500" = "Lawn, Garden, and Houseplants"
lawn_care <- atus_act |>
  filter(level2_code == "020500") |>
  inner_join(
    atus_resp |> select(participant_id, survey_weight,
                        is_employed, is_retired),
    join_by(participant_id)) |>
  mutate(status_group = case_when(
    is_retired  ~ "Retired",
    is_employed ~ "Employed",
    TRUE        ~ "Other / Not in Labor Force")) |>
  group_by(participant_id, status_group) |>
  summarize(daily_min    = sum(time_spent_min, na.rm = TRUE),
            survey_weight = first(survey_weight), .groups = "drop") |>
  group_by(status_group) |>
  summarize(avg_min_day  = weighted.mean(daily_min, survey_weight),
            avg_hrs_week = avg_min_day * 7 / 60,
            n            = n(), .groups = "drop") |>
  arrange(desc(avg_min_day))

lawn_care |>
  gt() |>
  tab_header(title    = "Time Spent on Lawn & Garden Activities",
             subtitle = "Weighted daily averages by employment status") |>
  cols_label(status_group = "Status", avg_min_day = "Avg Min/Day",
             avg_hrs_week = "Avg Hrs/Week", n = "Respondents") |>
  fmt_number(columns = c(avg_min_day, avg_hrs_week), decimals = 1) |>
  fmt_number(columns = n, decimals = 0) |>
  tab_style(
    style     = list(cell_fill(color = "#eaf4fb"), cell_text(weight = "bold")),
    locations = cells_body(rows = status_group == "Retired")) |>
  tab_source_note("Source: ATUS 2003-2024 (population-weighted)")
```

Retired Americans dedicate substantially more time to lawn and garden care. With no work hours to fill, retirees have the time and, typically, the homeownership situations that make sustained outdoor upkeep feasible.

#### 👫 Social Activities: Married vs. Unmarried

```{r table-social}
# code_level1 "120000" = Socializing, Relaxing, and Leisure
social_time <- atus_act |>
  filter(level1_code == "120000") |>
  inner_join(
    atus_resp |> select(participant_id, survey_weight, is_married),
    join_by(participant_id)) |>
  group_by(participant_id, is_married) |>
  summarize(daily_min    = sum(time_spent_min, na.rm = TRUE),
            survey_weight = first(survey_weight), .groups = "drop") |>
  group_by(is_married) |>
  summarize(avg_min_day  = weighted.mean(daily_min, survey_weight),
            avg_hrs_week = avg_min_day * 7 / 60,
            n            = n(), .groups = "drop") |>
  mutate(label = if_else(is_married, "Married (Spouse Present)", "Not Married"))

social_time |>
  select(label, avg_min_day, avg_hrs_week, n) |>
  gt() |>
  tab_header(title    = "Social & Leisure Time by Marital Status",
             subtitle = "Weighted average daily minutes") |>
  cols_label(label = "Status", avg_min_day = "Avg Min/Day",
             avg_hrs_week = "Avg Hrs/Week", n = "Respondents") |>
  fmt_number(columns = c(avg_min_day, avg_hrs_week), decimals = 1) |>
  fmt_number(columns = n, decimals = 0) |>
  tab_source_note("Source: ATUS 2003-2024 (population-weighted)")
```

Unmarried Americans report more social and leisure time, consistent with married individuals facing greater time constraints from household and caregiving obligations.

#### 💰 High-Earner Time Use

```{r table-high-income}
# is_high_income = annual_earn_est >= $100k (TRERNWA-based)
high_income_time <- atus_act |>
  inner_join(
    atus_resp |> filter(is_high_income) |> select(participant_id, survey_weight),
    join_by(participant_id)) |>
  inner_join(codes_l1, join_by(level1_code == code_level1)) |>
  group_by(participant_id, task_level1) |>
  summarize(daily_min    = sum(time_spent_min, na.rm = TRUE),
            survey_weight = first(survey_weight), .groups = "drop") |>
  group_by(task_level1) |>
  summarize(avg_min_day = weighted.mean(daily_min, survey_weight),
            .groups = "drop") |>
  arrange(desc(avg_min_day)) |>
  slice_head(n = 8)

high_income_time |>
  mutate(avg_hrs = avg_min_day / 60) |>
  gt() |>
  tab_header(title    = "Top Time Uses for High-Income Americans",
             subtitle = "Respondents with estimated annual earnings >= $100,000") |>
  cols_label(task_level1 = "Activity", avg_min_day = "Avg Min/Day",
             avg_hrs = "Avg Hrs/Day") |>
  fmt_number(columns = c(avg_min_day, avg_hrs), decimals = 1) |>
  data_color(columns = avg_min_day, palette = c("#eaf4fb","#1a5276")) |>
  tab_source_note("Income estimated from TRERNWA (weekly earnings x 52)")
```

Even for high earners, sleep and work dominate — a sobering reminder that money cannot buy more hours.

#### 🚌 Activities on Transit

```{r table-transit}
# Bus=15, Subway/Train=16, Boat/Ferry=18 per BLS codebook
transit_activities <- atus_act |>
  filter(on_transit) |>
  inner_join(codes_l2, join_by(level2_code == code_level2)) |>
  count(task_level2, wt = time_spent_min, name = "total_min") |>
  arrange(desc(total_min)) |>
  slice_head(n = 10) |>
  mutate(total_hrs  = total_min / 60,
         pct_of_top = total_min / sum(total_min))

transit_activities |>
  select(task_level2, total_hrs, pct_of_top) |>
  gt() |>
  tab_header(title    = "Top Activities While on Public Transit",
             subtitle = "Bus, subway/train, and boat/ferry (TEWHERE = 15, 16, 18)") |>
  cols_label(task_level2 = "Activity", total_hrs = "Total Hours",
             pct_of_top = "% of Top 10") |>
  fmt_number(columns = total_hrs, decimals = 0) |>
  fmt_percent(columns = pct_of_top, decimals = 1) |>
  data_color(columns = pct_of_top, palette = c("#fef9f3","#e67e22")) |>
  tab_source_note("Source: ATUS 2003-2024")
```

Transit time is dominated by commuting-adjacent activities, but leisure reclaiming of otherwise lost minutes is well represented.

#### 🎯 My Four Favorite Activities by Age

My four favorite activities are **Reading**, **Running**, **Cooking**, and **Socializing**.

```{r table-fav-activities}
# Level-3 codes verified against atus_activity_codes.csv:
# 120312 = "Reading for personal interest"
# 130124 = "Running"
# 020201 = "Food and drink preparation"
# 120101 = "Socializing and communicating with others"
fav_codes  <- c("120312", "130124", "020201", "120101")
fav_labels <- c("120312" = "Reading",
                "130124" = "Running",
                "020201" = "Cooking",
                "120101" = "Socializing")

fav_time <- atus_act |>
  filter(level3_code %in% fav_codes) |>
  inner_join(atus_resp |> select(participant_id, survey_weight, age),
             join_by(participant_id)) |>
  filter(!is.na(age)) |>
  mutate(
    activity_name = recode(level3_code, !!!fav_labels),
    age_group = cut(age,
      breaks = c(14, 24, 34, 44, 54, 64, Inf),
      labels = c("15-24","25-34","35-44","45-54","55-64","65+"),
      right  = TRUE)) |>
  group_by(activity_name, age_group) |>
  summarize(avg_min = weighted.mean(time_spent_min, survey_weight, na.rm = TRUE),
            .groups = "drop") |>
  pivot_wider(names_from = age_group, values_from = avg_min)

fav_time |>
  gt() |>
  tab_header(title    = "Daily Minutes on My Four Favorite Activities",
             subtitle = "Population-weighted averages by age group") |>
  cols_label(activity_name = "Activity") |>
  fmt_number(columns = -activity_name, decimals = 1) |>
  data_color(columns = -activity_name,
             palette = c("#ffffff","#1a5276"), na_color = "#f5f5f5") |>
  tab_source_note("Source: ATUS 2003-2024")
```

Reading rises sharply with age while running falls off after the 35-44 group. Cooking stays relatively stable across the lifespan.

:::

---

### 3.3 Graphical Explorations {#graph-explorations}

<span class="task-badge">Task 6</span>

::: {.panel-tabset}

#### 🏠 Household Activities vs. Age

```{r plot-household-age, fig.width=9, fig.height=5}
# code_level1 "020000" = Household Activities
household_by_age <- atus_act |>
  filter(level1_code == "020000") |>
  inner_join(atus_resp |> select(participant_id, survey_weight, age, sex),
             join_by(participant_id)) |>
  filter(!is.na(age), !is.na(sex), age >= 15, age <= 80) |>
  mutate(age_group = cut(age, breaks = seq(15, 80, by = 5),
                         right = FALSE, include.lowest = TRUE)) |>
  group_by(age_group, sex) |>
  summarize(avg_min = weighted.mean(time_spent_min, survey_weight, na.rm = TRUE),
            .groups = "drop") |>
  mutate(age_mid = as.numeric(str_extract(as.character(age_group), "\\d+")) + 2.5)

ggplot(household_by_age, aes(x = age_mid, y = avg_min, color = sex, fill = sex)) +
  geom_ribbon(aes(ymin = avg_min * 0.9, ymax = avg_min * 1.1),
              alpha = 0.15, color = NA) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 2.5, shape = 21, color = "white", stroke = 1.5) +
  scale_color_manual(values = c("Female" = "#e67e22","Male" = "#1a5276"),
                     name = "Sex") +
  scale_fill_manual(values  = c("Female" = "#e67e22","Male" = "#1a5276"),
                    name = "Sex") +
  scale_x_continuous(breaks = seq(20, 80, by = 10)) +
  labs(title   = "Household Activity Time Increases with Age - Especially for Women",
       subtitle = "Population-weighted avg minutes per episode by age group",
       x = "Age", y = "Avg Minutes per Activity",
       caption = "Source: ATUS 2003-2024")
```

Household activity time rises with age for both sexes, but women consistently spend more time per episode — a gap widening in the 30s and 40s during peak child-rearing years.

#### ☀️ Summer Activities in Education

```{r plot-education-summer, fig.width=9, fig.height=6}
# Filter to education sector workers in summer months.
# is_summer is on resp — join to act via participant_id.
edu_summer <- atus_act |>
  inner_join(
    atus_resp |>
      filter(in_education_sector, is_summer) |>
      select(participant_id, survey_weight),
    join_by(participant_id)) |>
  inner_join(codes_l2, join_by(level2_code == code_level2)) |>
  group_by(task_level2) |>
  summarize(total_wt_min = sum(time_spent_min * survey_weight, na.rm = TRUE),
            .groups = "drop") |>
  arrange(desc(total_wt_min)) |>
  slice_head(n = 12) |>
  mutate(task_level2 = fct_reorder(task_level2, total_wt_min))

ggplot(edu_summer, aes(x = total_wt_min / 1e9, y = task_level2,
                       fill = total_wt_min)) +
  geom_col(show.legend = FALSE, width = 0.7) +
  geom_text(aes(label = round(total_wt_min / 1e9, 1)),
            hjust = -0.1, size = 3.2, color = "#2c3e50") +
  scale_fill_gradient(low = "#aed6f1", high = "#1a5276") +
  scale_x_continuous(expand = expansion(mult = c(0, 0.2))) +
  labs(title   = "Top Summer Activities for Education Sector Workers",
       subtitle = "June-August; TRDTIND1 = 15 (Educational Services)",
       x = "Population-Weighted Total Minutes (Billions)", y = NULL,
       caption = "Source: ATUS 2003-2024")
```

During summer months, education workers prioritize leisure and personal care — confirming genuine rest rather than continued work obligations.

#### 🎓 Educational Time by Age for College Students

```{r plot-college-age, fig.width=9, fig.height=5}
# code_level1 "060000" = Education
college_study <- atus_act |>
  filter(level1_code == "060000") |>
  inner_join(
    atus_resp |>
      filter(is_college_student) |>
      select(participant_id, survey_weight, age),
    join_by(participant_id)) |>
  filter(!is.na(age), age >= 18, age <= 55) |>
  mutate(age_group = cut(age,
    breaks = c(17, 22, 26, 30, 40, 55),
    labels = c("18-22\n(Traditional)","23-26","27-30","31-40","41-55"),
    right  = TRUE)) |>
  group_by(participant_id, age_group) |>
  summarize(daily_min    = sum(time_spent_min, na.rm = TRUE),
            survey_weight = first(survey_weight), .groups = "drop") |>
  group_by(age_group) |>
  summarize(avg_hrs = weighted.mean(daily_min / 60, survey_weight),
            se      = sd(daily_min / 60) / sqrt(n()), .groups = "drop")

ggplot(college_study, aes(x = age_group, y = avg_hrs, fill = age_group)) +
  geom_col(show.legend = FALSE, width = 0.65) +
  geom_errorbar(aes(ymin = avg_hrs - se, ymax = avg_hrs + se),
                width = 0.2, color = "#5d6d7e") +
  geom_text(aes(label = round(avg_hrs, 1)), vjust = -0.8,
            fontface = "bold", color = "#1a5276", size = 4) +
  scale_fill_manual(
    values = c("#aed6f1","#5dade2","#2e86c1","#1a5276","#0d2b3e")) +
  labs(title   = "Older College Students Invest More Time in Education",
       subtitle = "Avg daily hours on educational activities among ATUS enrollees",
       x = "Age Group", y = "Avg Hours / Day",
       caption = "Source: ATUS 2003-2024 | Error bars = +/- 1 SE")
```

Older students dedicate measurably more time to education. Students aged 41-55 spend nearly 1.5x the hours of traditional-age (18-22) students.

#### 🏅 Bump Plot: Sports by Age Group

```{r plot-bump-sports, fig.width=10, fig.height=7}
# Specific level-3 sports codes from atus_activity_codes.csv
sport_codes <- tribble(
  ~level3_code, ~sport,
  "130103", "Basketball",
  "130114", "Golf",
  "130116", "Hiking",
  "130124", "Running",
  "130126", "Soccer",
  "130131", "Walking (Exercise)",
  "130133", "Weightlifting",
  "130136", "Yoga"
)

sports_by_age <- atus_act |>
  inner_join(sport_codes, join_by(level3_code)) |>
  inner_join(atus_resp |> select(participant_id, survey_weight, age),
             join_by(participant_id)) |>
  filter(!is.na(age), age >= 15, age <= 75) |>
  mutate(age_group = cut(age,
    breaks = c(14, 24, 34, 44, 54, 64, 75),
    labels = c("15-24","25-34","35-44","45-54","55-64","65-75"),
    right  = TRUE)) |>
  group_by(sport, age_group) |>
  summarize(total_wt = sum(time_spent_min * survey_weight, na.rm = TRUE),
            .groups = "drop") |>
  group_by(age_group) |>
  mutate(rank = rank(-total_wt, ties.method = "first")) |>
  ungroup() |>
  filter(rank <= 6)

ggplot(sports_by_age,
       aes(x = age_group, y = rank, group = sport, color = sport)) +
  geom_line(linewidth = 1.1, alpha = 0.85) +
  geom_point(size = 4.5, shape = 21, fill = "white", stroke = 2) +
  geom_text(data = sports_by_age |> filter(age_group == "15-24"),
            aes(label = sport, x = "15-24"),
            hjust = 1.05, size = 2.8, fontface = "italic") +
  scale_y_reverse(breaks = 1:6,
                  labels = c("1st","2nd","3rd","4th","5th","6th")) +
  scale_color_manual(values = scales::hue_pal()(12)) +
  scale_x_discrete(expand = expansion(add = c(2.5, 0.5))) +
  labs(title   = "Sports Rankings Shift Across the Lifespan",
       subtitle = "Top-6 activities by population-weighted total minutes, by age group",
       x = "Age Group", y = "Popularity Rank", color = NULL,
       caption = "Source: ATUS 2003-2024") +
  guides(color = guide_legend(nrow = 2))
```

Running and basketball dominate younger cohorts but give way to walking, golf, and yoga as Americans age — a natural shift toward lower-impact activity.

#### 👶 Child Care Per Child by Family Size

```{r plot-middle-child, fig.width=9, fig.height=5.5}
# code_level1 "030000" = Caring for and Helping Household Members
childcare_per_child <- atus_act |>
  filter(level1_code == "030000") |>
  inner_join(
    atus_resp |>
      filter(n_children >= 1, n_children <= 5) |>
      select(participant_id, survey_weight, n_children),
    join_by(participant_id)) |>
  group_by(participant_id, n_children) |>
  summarize(daily_min    = sum(time_spent_min, na.rm = TRUE),
            survey_weight = first(survey_weight), .groups = "drop") |>
  mutate(per_child_min = daily_min / n_children) |>
  group_by(n_children) |>
  summarize(avg_total    = weighted.mean(daily_min, survey_weight),
            avg_per_child = weighted.mean(per_child_min, survey_weight),
            .groups = "drop")

p1 <- ggplot(childcare_per_child,
             aes(x = factor(n_children), y = avg_total / 60)) +
  geom_col(fill = "#1a5276", width = 0.6) +
  geom_text(aes(label = round(avg_total / 60, 1)), vjust = -0.5,
            color = "#1a5276", fontface = "bold") +
  labs(title = "Total Childcare (hrs/day)",
       x = "# Children in Household", y = NULL) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)))

p2 <- ggplot(childcare_per_child,
             aes(x = factor(n_children), y = avg_per_child / 60)) +
  geom_col(fill = "#e67e22", width = 0.6) +
  geom_text(aes(label = round(avg_per_child / 60, 2)), vjust = -0.5,
            color = "#e67e22", fontface = "bold") +
  labs(title = "Childcare Per Child (hrs/day)",
       x = "# Children in Household", y = NULL) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)))

(p1 | p2) +
  plot_annotation(
    title   = "Middle-Child Syndrome: Care Per Child Falls as Family Size Grows",
    caption = "Source: ATUS 2003-2024")
```

As family size grows, total childcare increases — but not proportionally. A first child receives roughly twice the parental time a fourth or fifth child does.

:::

---

## 4 · Market Research: Target Demographics {#market-research}

<span class="task-badge">Task 7</span>

---

::: {.demo-card}
### 👨‍👧 Demographic 1: Stay-at-Home Dads

**Definition**: Males aged 25–55, not in the labor force, married (spouse present), with at least one child under 18.

Stay-at-home dads represent a growing demographic (~7% of fathers with children under 18).
:::

```{r sahd-profile}
sahd_ids <- atus_resp |>
  filter(
    sex == "Male",
    age %in% 25:55,
    employment_status == "Not in Labor Force",
    is_married,
    n_children >= 1
  ) |>
  select(participant_id, survey_weight, n_children)

sahd_act <- atus_act |>
  inner_join(sahd_ids, join_by(participant_id)) |>
  inner_join(codes_l1, join_by(level1_code == code_level1)) |>
  group_by(participant_id, task_level1) |>
  summarize(daily_min    = sum(time_spent_min, na.rm = TRUE),
            survey_weight = first(survey_weight), .groups = "drop") |>
  group_by(task_level1) |>
  summarize(low    = quantile(daily_min, 0.25),
            medium = weighted.mean(daily_min, survey_weight),
            high   = quantile(daily_min, 0.75), .groups = "drop") |>
  filter(task_level1 %in% c(
    "Caring for and Helping Household Members",
    "Household Activities",
    "Personal Care",
    "Eating and Drinking",
    "Sports, Exercise, and Recreation",
    "Socializing, Relaxing, and Leisure"
  )) |>
  mutate(
    target_min = case_match(task_level1,
      "Caring for and Helping Household Members" ~ 180,
      "Household Activities"                     ~ 120,
      "Personal Care"                            ~ 480,
      "Eating and Drinking"                      ~ 60,
      "Sports, Exercise, and Recreation"         ~ 45,
      "Socializing, Relaxing, and Leisure"       ~ 90,
      .default = 60),
    short_label = case_match(task_level1,
      "Caring for and Helping Household Members" ~ "Child Care",
      "Household Activities"                     ~ "Housework",
      "Personal Care"                            ~ "Sleep & Grooming",
      "Eating and Drinking"                      ~ "Meals",
      "Sports, Exercise, and Recreation"         ~ "Exercise",
      "Socializing, Relaxing, and Leisure"       ~ "Leisure",
      .default = task_level1),
    pct_hit = pnorm(target_min, mean = medium,
                    sd = pmax((high - low) / 1.35, 1),
                    lower.tail = FALSE)
  )

sahd_act |>
  select(short_label, low, medium, high, target_min, pct_hit) |>
  arrange(desc(medium)) |>
  gt() |>
  tab_header(title    = "Stay-at-Home Dad Time Use Profile",
             subtitle = "Daily minutes — typical range and aspirational targets") |>
  cols_label(short_label = "Activity", low = "25th %ile",
             medium = "Typical", high = "75th %ile",
             target_min = "Target", pct_hit = "% Hitting Target") |>
  fmt_number(columns = c(low, medium, high, target_min), decimals = 0) |>
  fmt_percent(columns = pct_hit, decimals = 0) |>
  data_color(columns = pct_hit, palette = c("#e74c3c","#f39c12","#27ae60")) |>
  tab_source_note("Targets based on pediatric and wellness recommendations")
```

**Key insight**: Stay-at-home dads are positioned to be primary caregivers yet many fall short of recommended interactive play time. The app could gamify milestones — "You're in the top 30% of hands-on dads this week!"

---

::: {.demo-card}
### 🎓 Demographic 2: College Undergraduates

**Definition**: Respondents enrolled in college or university (`is_college_student == TRUE`), aged 18–26.
:::

```{r undergrad-profile}
undergrad_ids <- atus_resp |>
  filter(is_college_student, age %in% 18:26) |>
  select(participant_id, survey_weight)

undergrad_act <- atus_act |>
  inner_join(undergrad_ids, join_by(participant_id)) |>
  inner_join(codes_l1, join_by(level1_code == code_level1)) |>
  group_by(participant_id, task_level1) |>
  summarize(daily_min    = sum(time_spent_min, na.rm = TRUE),
            survey_weight = first(survey_weight), .groups = "drop") |>
  group_by(task_level1) |>
  summarize(low    = quantile(daily_min, 0.25),
            medium = weighted.mean(daily_min, survey_weight),
            high   = quantile(daily_min, 0.75), .groups = "drop") |>
  filter(task_level1 %in% c(
    "Education",
    "Personal Care",
    "Socializing, Relaxing, and Leisure",
    "Eating and Drinking",
    "Sports, Exercise, and Recreation",
    "Work and Work-Related Activities"
  )) |>
  mutate(
    target_min = case_match(task_level1,
      "Education"                          ~ 240,
      "Personal Care"                      ~ 480,
      "Socializing, Relaxing, and Leisure" ~ 90,
      "Eating and Drinking"                ~ 60,
      "Sports, Exercise, and Recreation"   ~ 45,
      "Work and Work-Related Activities"   ~ 60,
      .default = 60),
    short_label = case_match(task_level1,
      "Education"                          ~ "Studying & Class",
      "Personal Care"                      ~ "Sleep & Grooming",
      "Socializing, Relaxing, and Leisure" ~ "Socializing & Leisure",
      "Eating and Drinking"                ~ "Meals",
      "Sports, Exercise, and Recreation"   ~ "Exercise",
      "Work and Work-Related Activities"   ~ "Part-Time Work",
      .default = task_level1),
    pct_hit = pnorm(target_min, mean = medium,
                    sd = pmax((high - low) / 1.35, 1),
                    lower.tail = FALSE)
  )

undergrad_act |>
  select(short_label, low, medium, high, target_min, pct_hit) |>
  arrange(desc(medium)) |>
  gt() |>
  tab_header(title    = "College Undergraduate Time Use Profile",
             subtitle = "Daily minutes — typical range and aspirational targets") |>
  cols_label(short_label = "Activity", low = "25th %ile",
             medium = "Typical", high = "75th %ile",
             target_min = "Target", pct_hit = "% Hitting Target") |>
  fmt_number(columns = c(low, medium, high, target_min), decimals = 0) |>
  fmt_percent(columns = pct_hit, decimals = 0) |>
  data_color(columns = pct_hit, palette = c("#e74c3c","#f39c12","#27ae60")) |>
  tab_source_note("Study target based on 15-credit course load recommendation")
```

**Key insight**: The gap between typical and target study time is the biggest app opportunity. Nudges like "20 more minutes puts you ahead of 60% of peers" could meaningfully shift behavior.

---

::: {.demo-card}
### 🏙️ Demographic 3: Young Professionals (25–35, Full-Time Employed, No Kids)

**Definition**: Respondents aged 25–35, employed full-time (`is_full_time == TRUE`), no children in household.
:::

```{r yuppie-profile}
yup_ids <- atus_resp |>
  filter(age %in% 25:35, is_full_time, n_children == 0) |>
  select(participant_id, survey_weight)

yup_act <- atus_act |>
  inner_join(yup_ids, join_by(participant_id)) |>
  inner_join(codes_l1, join_by(level1_code == code_level1)) |>
  group_by(participant_id, task_level1) |>
  summarize(daily_min    = sum(time_spent_min, na.rm = TRUE),
            survey_weight = first(survey_weight), .groups = "drop") |>
  group_by(task_level1) |>
  summarize(low    = quantile(daily_min, 0.25),
            medium = weighted.mean(daily_min, survey_weight),
            high   = quantile(daily_min, 0.75), .groups = "drop") |>
  filter(task_level1 %in% c(
    "Work and Work-Related Activities",
    "Personal Care",
    "Socializing, Relaxing, and Leisure",
    "Sports, Exercise, and Recreation",
    "Eating and Drinking",
    "Household Activities",
    "Volunteer Activities"
  )) |>
  mutate(
    target_min = case_match(task_level1,
      "Work and Work-Related Activities"   ~ 480,
      "Personal Care"                      ~ 450,
      "Socializing, Relaxing, and Leisure" ~ 90,
      "Sports, Exercise, and Recreation"   ~ 60,
      "Eating and Drinking"                ~ 60,
      "Household Activities"               ~ 45,
      "Volunteer Activities"               ~ 30,
      .default = 60),
    short_label = case_match(task_level1,
      "Work and Work-Related Activities"   ~ "Work",
      "Personal Care"                      ~ "Sleep & Grooming",
      "Socializing, Relaxing, and Leisure" ~ "Socializing & Leisure",
      "Sports, Exercise, and Recreation"   ~ "Exercise",
      "Eating and Drinking"                ~ "Meals",
      "Household Activities"               ~ "Chores",
      "Volunteer Activities"               ~ "Volunteering",
      .default = task_level1),
    pct_hit = pnorm(target_min, mean = medium,
                    sd = pmax((high - low) / 1.35, 1),
                    lower.tail = FALSE)
  )

yup_act |>
  select(short_label, low, medium, high, target_min, pct_hit) |>
  arrange(desc(medium)) |>
  gt() |>
  tab_header(title    = "Young Professional Time Use Profile",
             subtitle = "Ages 25-35, full-time employed, no children") |>
  cols_label(short_label = "Activity", low = "25th %ile",
             medium = "Typical", high = "75th %ile",
             target_min = "Target", pct_hit = "% Hitting Target") |>
  fmt_number(columns = c(low, medium, high, target_min), decimals = 0) |>
  fmt_percent(columns = pct_hit, decimals = 0) |>
  data_color(columns = pct_hit, palette = c("#e74c3c","#f39c12","#27ae60")) |>
  tab_source_note("Targets per WHO physical activity guidelines and burnout research")
```

```{r yup-visualization, fig.width=9, fig.height=5}
yup_chart <- yup_act |>
  mutate(
    short_label = fct_reorder(short_label, pct_hit),
    status = case_when(
      pct_hit >= 0.5 ~ "On Track",
      pct_hit >= 0.3 ~ "Near Miss",
      TRUE           ~ "Needs Work"))

ggplot(yup_chart, aes(x = pct_hit, y = short_label, color = status)) +
  geom_segment(aes(x = 0, xend = pct_hit, yend = short_label),
               linewidth = 1.2, alpha = 0.6) +
  geom_point(size = 5) +
  geom_vline(xintercept = 0.5, linetype = "dashed",
             color = "#27ae60", linewidth = 0.8) +
  annotate("text", x = 0.52, y = 1, label = "50% hit target",
           color = "#27ae60", size = 3, hjust = 0) +
  scale_x_continuous(labels = percent_format(), limits = c(0, 1),
                     expand = expansion(mult = c(0, 0.05))) +
  scale_color_manual(
    values = c("On Track" = "#27ae60","Near Miss" = "#f39c12",
               "Needs Work" = "#e74c3c"), name = NULL) +
  labs(title   = "Young Professionals: How Many Hit Their Targets?",
       subtitle = "% meeting or exceeding aspirational daily time allocations",
       x = "% of Population Meeting Target", y = NULL,
       caption = "Source: ATUS 2003-2024")
```

**Key insight**: Young professionals chronically undersleep and underexercise while overworking. Peer comparisons framed around "what peak performers in your bracket do" can drive strong engagement.

---

## 5 · Conclusion

This analysis of the American Time Use Survey reveals consistent patterns across 21 years:

1. **Sleep is non-negotiable** (~8.8 hrs/day), stable across demographics, and the single largest time expenditure.
2. **Household responsibilities distribute unevenly** by age, family size, and sex.
3. **Older students outperform** younger peers in academic dedication — a strong signal for adult learner marketing.
4. **Middle-child time dilution** is real and quantifiable.
5. **Young professionals** are the most time-pressured and data-hungry — a natural fit for peer-comparison gamification.

---

::: {.callout-note title="AI Usage Statement" collapse="true"}

l
:::

---

*Report generated using Quarto · Data: BLS ATUS 2003-2024 · 
